### 2021-05-18 proposal/slides.Rmd
updated slides

---
title: Community Detection Methods for the Generalized Random Dot Product Graph Model
subtitle: Dissertation Proposal Defense
author: John Koo
date: 'TBD'
output: 
  beamer_presentation:
    fig_crop: no
    theme: 'default'
    colortheme: 'beaver'
    includes:
      in_header: page_headers.tex
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE,
                      fig.align = 'center',
                      fig.lp = '')
options(xtable.comment = FALSE, 
        xtable.table.placement = 'H')
library(ggplot2)
import::from(magrittr, `%>%`)
theme_set(theme_bw())

source('~/dev/pabm-grdpg/functions.R')
set.seed(314159)
```

## Overview

* Preliminaries
  * Block Models and Community Detection
  * (Generalized) Random Dot Product Graphs
  * Connecting Block Models to the (G)RDPG
* The Popularity Adjusted Block Model
  * Connecting the PABM to the GRDPG
  * Subspace Clustering
* Community Detection for the (G)RDPG

# Preliminaries

## Bernoulli Graphs

<style type="text/css">
.caption {
    font-size: x-small;
}
</style>

::: columns

:::: {.column width=67%}

$A \sim BernoulliGraph(P)$ iff

* $G = (V, E)$
  * Undirected, unweighted, no self-loops
  * $|V| = n$
  * $|E| \leq n (n - 1) / 2$
* Adjacency matrix $A \in \{0, 1\}^{n \times n}$
  * $A_{ij} = \begin{cases} 
  1 & \exists \text{ edge between } i \text{ and } j \\
  0 & \text{else}
  \end{cases}$
* Edge probability matrix $P \in [0, 1]^{n \times n}$
* $A_{ij} \stackrel{indep}{\sim} Bernoulli(P_{ij})$ for $i < j$

::::

:::: {.column width=33%}

```{r, fig.height = 2, fig.width = 1.5}
n <- 2 ** 5
p <- 1 / log(n)
P <- matrix(p, nrow = n, ncol = n)
A <- draw.graph(P)
qgraph::qgraph(A, vsize = 4)
```

::::

:::

## Block Models

* $A \sim BernoulliGraph(P)$
  * (Hidden) labels $z_1, ..., z_n \in [K]$
  * $P_{ij} = f(z_i, z_j, \cdot)$

* Example: Stochastic Block Model with two communities
    * $K = 2$
  * $P_{ij} = \begin{cases} 
  p & z_i = z_j = 1 \\
  q & z_i = z_j = 2 \\
  r & z_i \neq z_j
  \end{cases}$

```{r, fig.height = 1.5, fig.width = 4}
n1 <- 2 ** 5
n2 <- 2 ** 5
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))
p <- 1/2
q <- 1/4
r <- 1/8
P <- matrix(r, nrow = n, ncol = n)
P[seq(n1), seq(n1)] <- p
P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
A <- draw.graph(P)
qgraph::qgraph(A, vsize = 4, groups = factor(z))
```

## Block Models

* Erdos-Renyi Model (1959)
  * $P_{ij} = \theta$
  * Not a block model
* Stochastic Block Model (Lorrain and White, 1971)
  * $P_{ij} = \theta_{z_i z_j}$
  * $K (K + 1) / 2$ parameters $\theta_{kl}$
* Degree Corrected Block Model (Karrer and Newman, 2011)
  * $P_{ij} = \theta_{z_i z_j} \omega_i \omega_j$
  * $K (K + 1) / 2 + n$ parameters $\theta_{kl}$, $\omega_i$
* Popularity Adjusted Block Model (Sengupta and Chen, 2017)
  * $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$
  * $K n$ parameters $\lambda_{ik}$

## Hierarchy of Block Models

* PABM $\to$ DCBM: $\lambda_{ik} = \sqrt{\theta_{z_i k}} \omega_i$
* DCBM $\to$ SBM: $\omega_i = 1$
* SBM $\to$ Erdos-Renyi: $\theta_{kl} = \theta$

\begin{center}
```{r, out.width = '150px'}
knitr::include_graphics('noroozi-pensky-hierarchy.png')
```

\tiny{Majid Noroozi and Marianna Pensky. "The Hierarchy of Block Models". March 2021.}
\end{center}

## Community Detection in Block Models

Likelihood

$$L = \prod_{i<j} \prod_{k, l}^K 
\big(p_{k, l, i, j}^{A_{ij}} 
(1 - p_{k, l, i, j})^{1 - A_{ij}} \big)^{z_{ik} z_{jl}}$$

* Example: DCBM ($p_{k, l, i, j} = \theta_{kl} \omega_i \omega_j$)

$$L = \prod_{i<j} \prod_{k, l}^K 
\big((\theta_{kl} \omega_i \omega_j)^{A_{ij}} 
(1 - \theta_{kl} \omega_i \omega_j)^{1 - A_{ij}} \big)^{z_{ik} z_{jl}}$$

* ML method for community detection: $\hat{\vec{z}} = \arg\max_{\vec{z}} L$

* NP-complete
  * Expectation-Maximization
  * Bayesian methods
  * **Spectral methods**

## (Generalized) Random Dot Product Graph Model

* Random Dot Product Graph $A \sim RDPG(X)$
  * Latent vectors $x_1, ..., x_n \in \mathcal{X} \subset \mathbb{R}^d$
    * $\mathcal{X} = \{x, y : 0 \leq x^\top y \leq 1\}$
  * Data matrix $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$
  * Edge probability matrix $P = X X^\top$
  * Adjacency matrix $A \sim BernoulliGraph(P)$

* Generalized Random Dot Product Graph $A \sim GRDPG_{p, q}(X)$
  * Latent vectors $x_1, ..., x_n \in \mathcal{X} \subset \mathbb{R}^d$
    * $\mathcal{X} = \{x, y : 0 \leq x^\top I_{p, q} y \leq 1\}$
    * $I_{p, q} = blockdiag(I_p, -I_q)$
    * $p + q = d$
  * Data matrix $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$
  * Edge probability matrix $P = X I_{pq} X^\top$
  * Adjacency matrix $A \sim BernoulliGraph(P)$

* If $X_1, ..., X_n \stackrel{iid}{\sim} F$, then $(A, X) \sim RDPG(F, n)$ or 
$(A, X) \sim GRDPG_{p, q}(F, n)$

## (Generalized) Random Dot Product Graph Model

### Recovery/Estimation

* Want to estimate $X$ given $A$
  * Alternatively, recover some property of $X$ given $A$ 
    * Interpoint distances
    * Inner products
    * Angles

* Adjacency Spectral Embedding
  * Given embedding dimension $d$, 
  $A \approx \hat{V} \hat{\Lambda} \hat{V}^\top$
    * If RDPG, use $d$ greatest eigenvalues
    * If GRDPG, use $p$ most positive and $q$ most negative eigenvalues
    * $\hat{V} \in \mathbb{R}^{n \times d}$
    * $\hat{\Lambda} \in \mathbb{R}^{d \times d}$
  * RDPG: $\hat{X} = \hat{V} \hat{\Lambda}^{1/2}$
  * GRDPG: $\hat{X} = \hat{V} |\hat{\Lambda}|^{1/2}$

* RDPG: $\max\limits_i \|\hat{X}_i - W_n X_i \| \stackrel{a.s.}{\to} 0$
(Athreya et al.)
* GRDPG: 
$\max\limits_i \|\hat{X}_i - Q_n X_i \| \stackrel{a.s.}{\to} 0$
(Rubin-Delanchy et al.)

## Connecting Block Models to the (G)RDPG Model

* All $G$ with $A \sim BernoulliGraph(P)$ are RDPG (if $P$ is 
positive semidefinite) or GRDPG
  * Includes all block models

* Example: SBM

$$X = \begin{bmatrix} 
\sqrt{p} & 0 \\
\vdots & \vdots \\
\sqrt{p} & 0 \\
\sqrt{r^2 / p} & \sqrt{q - r^2 / p} \\
\vdots & \vdots \\
\sqrt{r^2 / p} & \sqrt{q - r^2 / p}
\end{bmatrix}$$

$$P = X X^\top$$

## Connecting Block Models to the (G)RDPG Model

Example: SBM (cont'd)

* $A \sim BernoulliGraph(X X^\top)$
* $A \approx \hat{X} \hat{X}^\top$
  * $\hat{X} = \hat{V} \hat{\Lambda}^{1/2}$
* Apply clustering algorithm (e.g., $K$-means) on $\hat{X}$

```{r, fig.height = 8, fig.width = 8, out.width = '50%'}
A.eigen <- eigen(A)
X.hat <- A.eigen$vectors[, 1:2] %*% diag(A.eigen$values[1:2] ** .5)
plot(X.hat, asp = 1, col = z, xlab = NA, ylab = NA)
```

## Connecting Block Models to the (G)RDPG Model

```{r, out.width = '20%', fig.height = 2.5, fig.width = 2.5}
par(mar = rep(1.75, 4))
P.eigen = eigen(P)
X <- P.eigen$vectors[, 1:2] %*% diag(P.eigen$values[1:2] ** .5)
plot(X, asp = 1, col = z, xlab = NA, ylab = NA, main = 'SBM: Point masses')
```

```{r, out.width = '20%', fig.height = 2.5, fig.width = 2.5}
par(mar = rep(1.75, 4))
omega <- rbeta(n, 1, 1)
dc.matrix <- omega %*% t(omega)
P.dcbm <- P * dc.matrix
dcbm.eigen <- eigen(P.dcbm)
X.dcbm <- dcbm.eigen$vectors[, 1:2] %*% diag(dcbm.eigen$values[1:2] ** .5)
plot(X.dcbm, asp = 1, col = z, xlab = NA, ylab = NA, main = 'DCBM: Rays')

```

```{r, out.width = '33%', fig.height = 4, fig.width = 4}
par(mar = rep(1.75, 4))
Pz <- generate.P.beta(n)
P <- Pz$P
z <- Pz$clustering
X <- embedding(P)
pairs(X, col = z, asp = 1, pch = '.', main = 'PABM: Orthogonal subspaces')
```

# Popularity Adjusted Block Model

## Popularity Adjusted Block Model (Reparameterization)

$A \sim PABM(\{\lambda^{(kl)}\}_K)$ iff

* w.l.o.g., organize $P$ such that each block 
$P^{(kl)} \in [0, 1]^{n_k \times n_l}$ contains edge 
probabilities between communities $k$ and $l$
* Popularity vectors $\lambda^{(kl)} \in \mathbb{R}^{n_k}$ 
are the popularity parameters of members of community $k$ to community $l$
* $\{\lambda^{(kl)}\}_K$ is the set of $K^2$ popularity vectors
* $P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$
* $A \sim BernoulliGraph(P)$

## Connecting the PABM to the GRDPG ($K = 2$)

Theorem: $A \sim PABM(\{\lambda^{(kl)}\}_2)$ is equivalent to 
$A \sim GRDPG_{3, 1}(X U)$ for $X$ constructed from $\{\lambda^{(kl)}\}_2$ and 
$U \in \mathbb{O}(4)$

Proof:

$$X = \begin{bmatrix}
\lambda^{(11)} & \lambda^{(12)} & 0 & 0 \\
0 & 0 & \lambda^{(21)} & \lambda^{(22)}
\end{bmatrix}$$

$$Y = \begin{bmatrix}
\lambda^{(11)} & 0 & \lambda^{(12)} & 0 \\
0 & \lambda^{(21)} & 0 & \lambda^{(22)}
\end{bmatrix}$$

$$P = X Y^\top$$

## Connecting the PABM to the GRDPG ($K = 2$)

Proof (cont'd):

$$Y = X \Pi$$

$$\Pi = \begin{bmatrix} 
1 & 0 & 0 & 0 \\ 
0 & 0 & 1 & 0 \\
0 & 1 & 0 & 0 \\
0 & 0 & 0 & 1
\end{bmatrix} = 
U I_{3, 1} U^\top$$

$$U = \begin{bmatrix}
1 & 0 & 0 & 0 \\
0 & 0 & 1 / \sqrt{2} & 1 / \sqrt{2} \\
0 & 0 & 1 / \sqrt{2} & -1 / \sqrt{2} \\
0 & 1 & 0 & 0
\end{bmatrix}$$

$$P = (X U) I_{3, 1} (X U)^\top$$

## Connecting the PABM to the GRDPG

**Theorem**: $A \sim PABM(\{\lambda^{(kl)}\}_K)$ is equivalent to 
$A \sim GRDPG_{p, q}(X U)$ such that

* $p = K (K + 1) / 2$
* $q = K (K - 1) / 2$
* $U$ is orthogonal and predetermined for each $K$
* $X$ is block diagonal and composed of $\{\lambda^{(kl)}\}_K$
* $\implies$ if $x_i^\top$ and $x_j^\top$ are two rows of $X U$ corresponding 
to different communities, $x_i^\top x_j = 0$.

**Remark**: Non-uniqueness of the latent configuration  
$A \sim GRDPG_{p, q}(X U) \implies A \sim GRDPG_{p, q}(X U Q)$ 
$\forall Q \in \mathbb{O}(p, q)$

## Orthogonal Spectral Clustering

**Theorem**: If $P = V \Lambda V^\top$ and $B = n V V^\top$, then $B_{ij} = 0$ 
$\forall i, j$ in different communities.

Orthogonalized Spectral Clustering algorithm:

1. Compute the eigenvectors of $A$ that correspond to the $K (K+1)/2$ most 
positive and $K (K-1) / 2$ most negative eigenvalues to construct $V$.
2. Compute $B = |n V V^\top|$ applying $|\cdot|$ entry-wise.
3. Construct graph $G$ using $B$ as its similarity matrix.
4. Partition $G$ into $K$ disconnected subgraphs (e.g., using edge thresholding 
or spectral clustering).
5. Map each partition to the community labels $1, ..., K$.

## Orthogonal Spectral Clustering

**Theorem**: Let $\hat{B}_n$ with entries $\hat{B}_n^{(ij)}$ be the affinity 
matrix from OSC. 
Then $\forall$ pairs $(i, j)$ belonging to different communities 
and sparsity factor satisfying $n \rho_n = \omega\{(\log n)^{4c}\}$,

$$
\max_{i, j} \hat{B}^{(ij)} = 
O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)
$$

## Sparse Subspace Clustering

* $X$ is block diagonal and $U$ is orthogonal $\implies$ each community 
corresponds to a subspace in $\mathbb{R}^d$.
* Subspace property holds even with linear transformation 
$Q \in \mathbb{O}(p, q)$.
* If $P = V \Lambda V^\top$, then $V$ consists of *orthogonal* subspaces.

Sparse Subspace Clustering algorithm:

1. Solve $n$ optimization problems $c_i = \arg\min_c \|c\|_1$ 
subject to $x_i = X_{-i} c$ and $c_i^{(i)} = 0$.
2. Compile solutions $C = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix}$
3. Construct affinity matrix $B = |C| + |C^\top|$

* If $X$ obeys the Subspace Detection Property, then $B$ is sparse such that 
$B_{ij} = 0$ if $i$ and $j$ belong to different communities and $\|c_i\| > 0$.
* Step (1) of SSC typically performed via LASSO: 
$c_i = \arg\min \frac{1}{2} \|x_i - X_{-i} c\|_2^2 + \lambda \|c\|_1$

## Sparse Subspace Clustering

**Theorem**:

Let

* $P_n$ describe the edge probability matrix of the PABM with 
$n$ vertices
* $A_n \sim BernoulliGraph(P_n)$
* $\hat{V}_n$ be the matrix of eigenvectors of $A_n$ corresponding to the 
$K (K + 1) / 2$ most positive and $K (K - 1) / 2$ most negative eigenvalues.

Then

* $\exists \lambda > 0$ and $N < \infty$ such that when $n > N$, 
$\sqrt{n} \hat{V}_n$ obeys the Subspace Detection Property with probability 1.

Remarks:

* For large $n$, we can identify $\lambda$ for SDP (Wang and Xu, 2016).
* SDP does not guarantee community detection.

# General Community Detection for the (G)RDPG

## Generative Model

Let $(A, X) \sim RDPG(F, n)$ such that

1. Define functions $f_1, ..., f_K$ such that 
$f_k : [0, 1] \mapsto \mathcal{X}$ and $f_k(t) \neq f_l(t)$ 
$\forall k, l \in [K]$. 
2. Sample labels
$Z_1, ..., Z_n \stackrel{iid}{\sim} Categorical(\pi_1, ..., \pi_K)$.
3. Sample $T_1, ..., T_n \stackrel{iid}{\sim} D$ with support $[0, 1]$. 
4. Set latent positions $X_i = f_{Z_i}(T_i)$ and 
$X = \begin{bmatrix} X_1 & \cdots & X_n \end{bmatrix}^\top$.
5. $A \sim BernoulliGraph(X X^\top)$

## Community Detection

* Athreya et al. and Rubin-Delanchy et al.: we can approximate properties of 
the latent configurations via ASE.
* General community detection method: Given $A$, $K$, and $d$ (or $p$ and $q$), 
  1. Use ASE to approximate the latent configuration.
  2. Use the appropriate clustering algorithm for the form of the latent
  configuration (manifolds).

## Parallel Segments

**Example**: 
Let $U_1, ..., U_{n_1}, U_{n_1 + 1}, ..., U_{n} \stackrel{iid}{\sim} Uniform(0, \cos \frac{\pi}{2} a)$, $f_1(t) = (t, 0)$, and $f_2(t) = (t, a)$. $X_i = f_1(U_i)$ for $i \leq n_1$ and $X_j = f_2(U_j)$ for $n_1 + 1 \leq j \leq n$. If we observe $X_1, ..., X_{n_1}, X_{n_1+1}, ..., X_{n}$, what approach will allow us to group the observations by segment?

```{r fig.height = 3, out.width = '50%'}
set.seed(1)

n1 <- 2 ** 6
n2 <- 2 ** 6
n <- n1 + n2

a <- .6

X <- cbind(runif(n1, 0, cos(pi / 2 * a)), rep(0, n1))
Y <- cbind(runif(n2, 0, cos(pi / 2 * a)), rep(a, n2))

data.matrix <- rbind(X, Y)
z <- c(rep(1, n1), rep(2, n2))

as.data.frame(data.matrix) %>% 
  dplyr::mutate(label = factor(z)) %>% 
  ggplot() + 
  geom_segment(aes(x = 0, xend = cos(pi / 2 * a), y = 0, yend = 0), size = .1) + 
  geom_segment(aes(x = 0, xend = cos(pi / 2 * a), y = a, yend = a), size = .1) + 
  geom_point(aes(x = V1, y = V2, colour = label)) + 
  labs(x = NULL, y = NULL, colour = NULL) + 
  guides(colour = FALSE) + 
  xlim(-.1, cos(pi / 2 * a) + .1) + 
  ylim(-a, 2 * a) + 
  scale_colour_brewer(palette = 'Set1') + 
  # coord_fixed() + 
  # ggthemes::theme_few() + 
  theme_void() + 
  theme(axis.text = element_blank(),
        axis.ticks = element_blank())
```

$\forall a \in (0, 1)$, $\delta \in (0, 1)$, and $K \geq 2$, 
$\exists N(a, \delta, K) < \infty$ such that when $\min_k n_k \geq N$, 
with probability at least $1 - \delta$,

1. Single linkage clustering will produce perfect community detection.
2. Any $\epsilon$-neighborhood graph with $\epsilon \leq a$ will consist of 
at least $K$ disjoint subgraphs such that no subgraph contains members of two 
different communities.

## Noisy Parallel Segments and One-Dimensional Manifolds

**Example**: 
Starting with the parallel segments as before, suppose instead of observing $X_1, ..., X_n$, we have noisy observations $X_1 + \xi_1, ..., X_n + \xi_n$ such that 
$\max_i \|\xi_i\| = \xi \leq a / 3$.

```{r fig.height = 3, out.width = '50%'}
# xi1 <- runif(n, -a / 6 / sqrt(2), a / 6 / sqrt(2))
xi1 <- rep(0, n)
xi2 <- runif(n, -a / 6 / sqrt(2), a / 6 / sqrt(2))

as.data.frame(data.matrix) %>% 
  dplyr::mutate(V1 = V1 + xi1, V2 = V2 + xi2) %>% 
  dplyr::mutate(label = factor(z)) %>% 
  ggplot() + 
  geom_segment(aes(x = 0, xend = cos(pi / 2 * a), y = 0, yend = 0), size = .1) + 
  geom_segment(aes(x = 0, xend = cos(pi / 2 * a), y = a, yend = a), size = .1) + 
  geom_point(aes(x = V1, y = V2, colour = label)) + 
  labs(x = NULL, y = NULL, colour = NULL) + 
  guides(colour = FALSE) + 
  xlim(-.1, cos(pi / 2 * a) + .1) + 
  ylim(-a, 2 * a) + 
  scale_colour_brewer(palette = 'Set1') + 
  # coord_fixed() + 
  # ggthemes::theme_few() + 
  theme_void() + 
  theme(axis.text = element_blank(),
        axis.ticks = element_blank())
```

Then $\forall a \in (0, 1)$, $\delta \in (0, 1)$, $K \geq 2$, $\xi \leq a / 3$, 
$\exists N(a, \delta, K, \xi) < \infty$ such that when $\min_k n_k \geq N$, 
with probability at least $1 - \delta$,

1. Single linkage will produce perfect community detection.
2. Any $\epsilon$-neighborhood graph will consist of at least $K$ sub-graphs with no subgraph containing vertices from multiple communities.

This also holds for noisy points sampled from one-dimensional manifolds such that the manifolds are distance at least $a$ apart.

## Recovery from the Adjacency Matrix

::: columns

:::: {.column width=33%}

```{r, fig.height = 2, fig.width = 3, out.width = '100%'}
par(mar = rep(1, 4))
plot(data.matrix, col = z * 2, asp = 1, xlab = NA, ylab = NA,
     main = 'Latent configuration')
```

::::

:::: {.column width=33%}

```{r, fig.height = 2, fig.width = 3, out.width = '100%'}
par(mar = rep(1, 4))
P <- data.matrix %*% t(data.matrix)
A <- draw.graph(P)
qgraph::qgraph(A, groups = factor(z))
title('Sampled graph')
```

::::

:::: {.column width=33%}

```{r, fig.height = 2, fig.width = 3, out.width = '100%'}
par(mar = rep(1, 4))
A.eigen <- eigen(A)
Xhat <- A.eigen$vectors[, 1:2] %*% diag(sqrt(A.eigen$values[1:2]))
plot(Xhat, col = z * 2, asp = 1, main = 'ASE')
```

::::

:::

## Future Work

1. Show that the ASE of a random graph generated by these latent vectors 
produces the correct conditions for sufficiently large $n$.
2. Extend results to non-uniform distributions.
3. Extend results to multidimensional manifolds.
4. Relax condition for the minimum distance between manifolds.
5. Explore more robust clustering techniques for these latent configurations.
6. Extend results to the GRDPG.

# Additional Slides

## Indefinite Orthogonal Group

$$\mathbb{O}(p, q) = \{Q : Q I_{p, q} Q^\top = I_{p, q}\}$$

* $Q^\top Q \neq I$
* If $A \sim GRDPG_{p, q}(X)$, then $A \sim GRDPG_{p, q}(X Q)$
* $(Q x)^\top (Q y) = x^\top Q^\top Q y \neq x^\top y$
* $\|Q\| \neq 1$ $\implies$ $\|Q x - Q y\| \neq \|x - y\|$

## Expectation Maximization for the PABM

Full data log-likelihood

$$\begin{split}
\log L = & 
\sum_{i < j} \sum_{k, l} z_{ik} z_{jl} (A_{ij} \log \lambda_{il} \lambda_{jk} +
(1 - A_{ij}) \log(1 - \lambda_{il} \lambda_{jk})) \\
+ & \sum_i \sum_k z_{ik} \log \pi_k
\end{split}$$

* E-step
  * $\gamma_{ik} = P(Z_i = k \mid \{\pi_l\}, \{\lambda_{jl}\})$
  * $\log \gamma_{ik} \propto 
  \log \pi_k + 
  \sum_{j \neq i} \sum_l \pi_{jl} (A_{ij} \log \lambda_{il} \lambda_{jk} + 
  (1 - A_{ij}) \log(1 - \lambda_{il} \lambda_{jk}))$
* M-step
  * $\pi_k = \frac{1}{n} \sum_i \gamma_{ik}$
  * $\{\lambda_{ik}\} = \arg\max_{\{\lambda_{ik}\}} E_Z[\log L]$

## MCMC Sampling for the PABM

Priors:

* $Z_i \stackrel{iid}{\sim} Categorical(\pi_1, ..., \pi_K)$
* $\lambda_{ik} \stackrel{indep}{\sim} Beta(a_{ik}, b_{ik})$

Full joint distribution:

$$\begin{split}
\log p & = constant \\
& + \sum_{i < j} \sum_k \sum_l z_{ik} z_{jl} 
(A_{ij} \log \lambda_{il} \lambda_{jk} + 
(1 - A_{ij}) \log (1 - \lambda_{il} \lambda_{jk})) \\
& + \sum_k \sum_i z_{ik} \log \pi_k \\
& + \sum_i \sum_k (a_{ik} - 1) \log \lambda_{ik} + (b_{ik} - 1) \log (1 - \lambda_{ik})
\end{split}$$

## Variational Inference for the PABM

Mean Field Variational Inference

* Minimize $d_{KL}(p || q)$
  * $p$ is the joint distribution
  * $q$ is a density of some form
* Restrict 
$q(\vec{z}, \{\lambda_{ik}\}) = 
\Big(\prod_i q_{z_i}(z_i) \Big) \Big(\prod_{i, k} q_{\lambda_{ik}}(\lambda_{ik})\Big)$
* Iterative solution: 
$q_{\theta_i}^{(t+1)} \propto \exp(E_{\theta_{-i}^{(t)}}[\log p])$

* Approximate solution for the PABM
  * $Z_i \mid \{a'_{ik}\}, \{b'_{ik}\} \sim Categorical(\pi'_1, ..., \pi'_K)$
  * $\lambda_{ik} \mid 
  \{a'_{-i, -k}\}, \{b'_{-i, -k}\}, \{\pi'_k\} \sim Beta(a'_{ik}, b'_{ik})$
  * Iteratively update $\{\pi'_K\}, \{a'_{ik}\}, \{b'_{ik}\}$ until convergence