### 2021-06-14 proposal/prospectus.Rmd
updated prospectus

---
title: |
  | Community Detection for the 
  | Generalized Random Dot Product Graph
subtitle: Dissertation Prospectus
author: John Koo
institution: Indiana University
output:
  pdf_document:
    keep_tex: true
    citation_package: natbib
    number_sections: yes
fontsize: 11pt
# geometry: "left=1in,right=1in,top=1in,bottom=1in"
urlcolor: blue
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
bibliography: proposal.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      # eval = FALSE,
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.pos = 'H', 
                      fig.align = 'center', 
                      fig.height = 3, 
                      fig.width = 5, 
                      fig.dpi = 300)

options(xtable.comment = FALSE, 
        xtable.table.placement = 'H')

import::from(magrittr, `%>%`)
library(ggplot2)
```

\newcommand{\diag}{\text{diag}}
\newcommand{\tr}{\text{Tr}}
\newcommand{\blockdiag}{\text{blockdiag}}
\newcommand{\indep}{\stackrel{\text{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\text{iid}}{\sim}}
\newcommand{\Bernoulli}{\text{Bernoulli}}
\newcommand{\Betadist}{\text{Beta}}
\newcommand{\BG}{\text{BernoulliGraph}}
\newcommand{\Cat}{\text{Categorical}}
\newcommand{\GRDPG}{\text{GRDPG}}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\theoremstyle{example}
\newtheorem*{example}{Example}

Graph and network data, in which samples are represented not as a collection of feature vectors but as relationships between pairs of observations, are increasingly widespread in various fields ranging from sociology to computer vision. Analyzing such data, however, can be a challenge because traditional statistical and machine learning techniques are often incompatible with graph data. For example, if we observe data sampled from a mixture of Gaussians, we can iteratively update the means and covariances by computing summary statistics for each mixture component and update community labels via Mahalanobis distances using the sample means and covariances \cite{doi:10.1198/016214502760047131}. A type of model that is analogous to mixture models for graph data is the Block Model, which is a family of generative models that have community structure. For such models, it is often impossible to compute analogous summary statistics (e.g., sample mean) for community detection or parameter estimation. In this research, we investigate the use of a family of spectral decomposition and embedding based approaches for community detection and parameter estimation in block models (random graph models with inherent community structure). First, we demonstrate that under the Generalized Random Dot Product Graph \cite{rubindelanchy2017statistical} framework, all graphs generated by Block Models can be represented as collections of vectors in Euclidean space via the Adjacency Spectral Embedding \cite{athreya2017statistical, lyzinski2014}. Then, noting the particular structure or configuration that the vectors take, we select appropriate clustering and parameter estimation techniques to apply on the vector representation of the graph. Finally we derive the asymptotic properties of these methods to show that our methods are consistent or achieve desirable properties with high probability. To illustrate this approach, we primarily focus on a type of Block Model called the Popularity Adjusted Block Model \cite{307cbeb9b1be48299388437423d94bf1}.

Our work is essentially about connecting two well-known families of generative graph models, the Block Model and the (Generalized) Random Dot Product Graph \cite{rubindelanchy2017statistical} in order to perform statistical inference. It is straightforward to show that all Block Models are special cases of the GRDPG. This fact has been leveraged to develop community detection and parameter estimation methods \cite{athreya2017statistical, lyzinski2014, rubindelanchy2017statistical} for the Stochastic Block Model \cite{doi:10.1080/0022250X.1971.9989788} and Degree Corrected Block Model \cite{Karrer_2011}. Our work begins by extending this to the Popularity Adjusted Block Model. We show that under the GRDPG framework, the latent configuration that generates the PABM consists of orthogonal subspaces. The Adjacency Spectral Embedding allows us to recover this latent configuration from an observed adjacency matrix. Then we show that two community detection algorithms arise naturally from this latent configuration: Orthogonal Spectral Clustering, which is an algorithm of our own design, and an existing algorithm, Sparse Subspace Clustering \cite{5206547}. Parameter estimation via spectral decomposition can be performed in a similar vein. We then extend this work to more general configurations in the latent space that imply community structure. The overall goal of our research is to develop a general framework or approach to statistical inference for a wide range of generative graph models with community structure by investigating their latent structure under the GRDPG framework.

The overall generative model for inducing community structure can be described by the following.

Let $(A, X) \sim \GRDPG_{p, q}(F, n)$ such that:

1. Define functions $\gamma_1, ..., \gamma_K$ such that each 
$\gamma_k : [0, 1]^r \mapsto \mathbb{R}^d$ and $\gamma_k(t) \neq \gamma_l(t)$ 
when $k \neq l$.
2. Sample labels
$Z_1, ..., Z_n \iid \Cat(\pi_1, ..., \pi_K)$.
3. Sample $T_1, ..., T_n \iid D$ with support $[0, 1]^r$. 
4. Set latent positions $X_i = \gamma_{Z_i}(T_i)$ and 
$X = \begin{bmatrix} X_1 & \cdots & X_n \end{bmatrix}^\top$.
5. $A \sim \BG(X I_{p, q} X^\top)$

The SBM, DCBM, and PABM are all special cases of this generative model. Based on the structure of the $\gamma_k$ functions that correspond to each of the SBM, DCBM, and PABM, there are natural choices for clustering algorithms on the embeddings of each of these Block Models, followed by straightforward parameter estimation techniques. Extensions of these results may consider which clustering algorithms result in consistent estimators given restrictions on each of the $\gamma_k$, and how to approach community detection when $\gamma_k$ are unknown altogether.

The estimated timeline of completion is as follows:

1. Literature review and proofs of main theorems: December 2022

2. Simulations, real data analyses, R package: March 2022

3. Dissertation Completion: April 2022