### 2021-06-14 proposal/prospectus.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
  11pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdfauthor={John Koo},
  colorlinks=true,
  linkcolor=Maroon,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=blue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{setspace}
\usepackage{float}
\usepackage{mathtools}
\usepackage{natbib}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\setcitestyle{numbers,square,comma}
\usepackage{verbatim}
\usepackage{amsthm}
\usepackage{comment}
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{plainnat}

\title{Community Detection for the\\
Generalized Random Dot Product Graph}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Dissertation Prospectus}
\author{John Koo}
\date{}

\begin{document}
\maketitle

\newcommand{\diag}{\text{diag}}
\newcommand{\tr}{\text{Tr}}
\newcommand{\blockdiag}{\text{blockdiag}}
\newcommand{\indep}{\stackrel{\text{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\text{iid}}{\sim}}
\newcommand{\Bernoulli}{\text{Bernoulli}}
\newcommand{\Betadist}{\text{Beta}}
\newcommand{\BG}{\text{BernoulliGraph}}
\newcommand{\Cat}{\text{Categorical}}
\newcommand{\GRDPG}{\text{GRDPG}}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\theoremstyle{example}
\newtheorem*{example}{Example}

Graph and network data, in which samples are represented not as a
collection of feature vectors but as relationships between pairs of
observations, are increasingly widespread in various fields ranging from
sociology to computer vision. Analyzing such data, however, can be a
challenge because traditional statistical and machine learning
techniques are often incompatible with graph data. For example, if we
observe data sampled from a mixture of Gaussians, we can iteratively
update the means and covariances by computing summary statistics for
each mixture component and update community labels via Mahalanobis
distances using the sample means and covariances
\cite{doi:10.1198/016214502760047131}. A type of model that is analogous
to mixture models for graph data is the Block Model, which is a family
of generative models that have community structure. For such models, it
is often impossible to compute analogous summary statistics (e.g.,
sample mean) for community detection or parameter estimation. In this
research, we investigate the use of a family of spectral decomposition
and embedding based approaches for community detection and parameter
estimation in block models (random graph models with inherent community
structure). First, we demonstrate that under the Generalized Random Dot
Product Graph \cite{rubindelanchy2017statistical} framework, all graphs
generated by Block Models can be represented as collections of vectors
in Euclidean space via the Adjacency Spectral Embedding
\cite{athreya2017statistical, lyzinski2014}. Then, noting the particular
structure or configuration that the vectors take, we select appropriate
clustering and parameter estimation techniques to apply on the vector
representation of the graph. Finally we derive the asymptotic properties
of these methods to show that our methods are consistent or achieve
desirable properties with high probability. To illustrate this approach,
we primarily focus on a type of Block Model called the Popularity
Adjusted Block Model \cite{307cbeb9b1be48299388437423d94bf1}.

Our work is essentially about connecting two well-known families of
generative graph models, the Block Model and the (Generalized) Random
Dot Product Graph \cite{rubindelanchy2017statistical} in order to
perform statistical inference. It is straightforward to show that all
Block Models are special cases of the GRDPG. This fact has been
leveraged to develop community detection and parameter estimation
methods
\cite{athreya2017statistical, lyzinski2014, rubindelanchy2017statistical}
for the Stochastic Block Model \cite{doi:10.1080/0022250X.1971.9989788}
and Degree Corrected Block Model \cite{Karrer_2011}. Our work begins by
extending this to the Popularity Adjusted Block Model. We show that
under the GRDPG framework, the latent configuration that generates the
PABM consists of orthogonal subspaces. The Adjacency Spectral Embedding
allows us to recover this latent configuration from an observed
adjacency matrix. Then we show that two community detection algorithms
arise naturally from this latent configuration: Orthogonal Spectral
Clustering, which is an algorithm of our own design, and an existing
algorithm, Sparse Subspace Clustering \cite{5206547}. Parameter
estimation via spectral decomposition can be performed in a similar
vein. We then extend this work to more general configurations in the
latent space that imply community structure. The overall goal of our
research is to develop a general framework or approach to statistical
inference for a wide range of generative graph models with community
structure by investigating their latent structure under the GRDPG
framework.

The overall generative model for inducing community structure can be
described by the following.

Let \((A, X) \sim \text{GRDPG}_{p, q}(F, n)\) such that:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Define functions \(\gamma_1, ..., \gamma_K\) such that each
  \(\gamma_k : [0, 1]^r \mapsto \mathbb{R}^d\) and
  \(\gamma_k(t) \neq \gamma_l(t)\) when \(k \neq l\).
\item
  Sample labels
  \(Z_1, ..., Z_n \stackrel{\text{iid}}{\sim}\text{Categorical}(\pi_1, ..., \pi_K)\).
\item
  Sample \(T_1, ..., T_n \stackrel{\text{iid}}{\sim}D\) with support
  \([0, 1]^r\).
\item
  Set latent positions \(X_i = \gamma_{Z_i}(T_i)\) and
  \(X = \begin{bmatrix} X_1 & \cdots & X_n \end{bmatrix}^\top\).
\item
  \(A \sim \text{BernoulliGraph}(X I_{p, q} X^\top)\)
\end{enumerate}

The SBM, DCBM, and PABM are all special cases of this generative model.
Based on the structure of the \(\gamma_k\) functions that correspond to
each of the SBM, DCBM, and PABM, there are natural choices for
clustering algorithms on the embeddings of each of these Block Models,
followed by straightforward parameter estimation techniques. Extensions
of these results may consider which clustering algorithms result in
consistent estimators given restrictions on each of the \(\gamma_k\),
and how to approach community detection when \(\gamma_k\) are unknown
altogether.

The estimated timeline of completion is as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Literature review and proofs of main theorems: December 2022
\item
  Simulations, real data analyses, R package: March 2022
\item
  Dissertation Completion: April 2022
\end{enumerate}

  \bibliography{proposal.bib}

\end{document}