### 2021-08-18 scratch/distributions-on-1d-manifolds.Rmd
started typing up stuff

---
title: |
  | Clustering of Distributions on 1-Dimensional Manifolds
output:
  pdf_document:
    keep_tex: true
    citation_package: natbib
fontsize: 11pt
# geometry: "left=1in,right=1in,top=1in,bottom=1in"
urlcolor: blue
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
bibliography: proposal.bib
---

## Setup

Suppose we have two manifolds $\mathcal{M}_1$ and $\mathcal{M}_2$ $\in \mathbb{R}^d$, each of length 1, defined by $f_1(t)$ and $f_2(t)$ respectively ($f_i : [0, 1] \mapsto \mathbb{R}^d$). Define $\delta$ as the minimum distance between the two manifolds, i.e., $\delta = \min_t \|f_1(t) - f_2(t)\|$, and let $\delta > 0$. For now, restrict each $f_i$ such that the distance along the manifold between $f_i(t)$ and $f_i(s)$ is equal to the difference between $t$ and $s$ (this also implies that each manifold is of length 1). 
We sample $T_1, ..., T_n \stackrel{iid}{\sim} F$ for continuous $F$ with support $[0, 1]$ and use $f_1$ to map the first $n_1$ points to $\mathcal{M}_1$ and $f_2$ to map the remaining $n_2 = n - n_1$ points to $\mathcal{M}_2$. Let $X_i = f_1(T_i)$ and $Y_j = f_2(T_j)$.
Without loss of generality, assume $n_1 \leq n_2$.

## Theory

Let $D_i = X_{(i)} - X_{(i-1)}$. Then if $\max_i D_i < \delta$, we have sufficient separation of points in $\mathcal{M}_1$.

### Uniform case

It can be shown that if $X_i \stackrel{iid}{\sim} Uniform(0, 1)$, then $D_i \sim Beta(1, n)$. Therefore, $P(\max_i D_i < \delta) \geq (P(D_i < \delta))^n = (1 - (1 - \delta)^n)^n$, which is a decreasing function for sufficiently large $n$. This gives us the result $\max_i D_i \stackrel{a.s.}{\to} 0$.

### Beta case

### General case

If $F$ is absolutely continuous, then it can be shown that

$$P(D_i \leq \delta) = 1 - \int_0^{1-\delta} (i-1) \binom{n}{i-1} f(x) (F(x))^{i-1} (1 - F(x + \delta))^{n-i} dx$$

Under certain conditions (namely $F(x + \delta) \geq F(x) + \delta$), we can make the substitution $u = F(x)$ and have the result

$$P(D_i \leq \delta) \leq 1- \int_0^{1 - \delta} (i-1) \binom{n}{i-1} u^{i-1} (1 - u - \delta)^{n-i} du$$

which gives the same result as in the uniform case.

## Computational Results

TBD