### 2021-09-15 scratch/distributions-on-1d-manifolds.Rmd
cleaned up some stuff

---
title: |
  | Clustering of Distributions on 1-Dimensional Manifolds
output:
  pdf_document:
    keep_tex: true
    citation_package: natbib
fontsize: 11pt
# geometry: "left=1in,right=1in,top=1in,bottom=1in"
urlcolor: blue
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
bibliography: proposal.bib
---

\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{Tr}}
\newcommand{\blockdiag}{\mathrm{blockdiag}}
\newcommand{\indep}{\stackrel{\mathrm{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
\newcommand{\Bernoulli}{\mathrm{Bernoulli}}
\newcommand{\Betadist}{\mathrm{Beta}}
\newcommand{\Uniform}{\mathrm{Uniform}}
\newcommand{\BG}{\mathrm{BernoulliGraph}}
\newcommand{\Categorical}{\mathrm{Categorical}}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\theoremstyle{example}
\newtheorem*{example}{Example}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\as}{\stackrel{\mathrm{a.s.}}{\to}}

# Setup

Suppose we have two manifolds $\mathcal{M}_1$ and $\mathcal{M}_2$ $\in \mathbb{R}^d$, each of length 1, defined by $f_1(t)$ and $f_2(t)$ respectively ($f_i : [0, 1] \mapsto \mathbb{R}^d$). Define $\delta$ as the minimum distance between the two manifolds, i.e., $\delta = \min_t \|f_1(t) - f_2(t)\|$, and let $\delta > 0$. For now, restrict each $f_i$ such that the distance along the manifold between $f_i(t)$ and $f_i(s)$ is equal to the difference between $t$ and $s$ (this also implies that each manifold is of length 1). 
We sample $T_1, ..., T_n \stackrel{iid}{\sim} F$ for continuous $F$ with support $[0, 1]$ and use $f_1$ to map the first $n_1$ points to $\mathcal{M}_1$ and $f_2$ to map the remaining $n_2 = n - n_1$ points to $\mathcal{M}_2$. Let $X_i = f_1(T_i)$ and $Y_j = f_2(T_j)$.
Without loss of generality, assume $n_1 \leq n_2$.

# Preliminary Theory

## Distributions of differences of order statistics

Let $D_i = X_{(i+1)} - X_{(i)}$. Then if $\max_i D_i < \delta$, we have sufficient separation of points in $\mathcal{M}_1$. 
Then it is sufficient to quantify $P(\max_i D_i > \delta)$ as a function of $n$ and $\delta$ and show that this converges to zero as $n$ grows to $\infty$.

We denote $f(x)$ as the density of each $X_i$, $g_i(x)$ as the density of $X_{(i)}$, 
$g_{ij}(x, y)$ as the joint density of $X_{(i)}, X_{(j)}$, and $h_i(d)$ as the density of $D_i$ (with corresponding capital letters for the cumulative distribution functions).

The following are taken as given[^wikipedia]:

1. $g_i(x) = \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x))^{n-i} f(x)$.
2. $g_{ij}(x) = \frac{n!}{(i-1)! (j-i-1)! (n-j)!} (F(x))^{i-1} (F(y) - F(x))^{j-i-1} (1 - F(y))^{n-j} f(x) f(y)$.
3. By convolution, $h_i(d) = \int_0^{1} g_{i, i+1} (x, x + d) \dd x$.

[^wikipedia]: https://en.wikipedia.org/wiki/Order_statistic

\begin{lemma}[The probability density function of $D_i$]
\begin{equation}
\label{eq:pdf}
h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} (1 - F(x+d))^{n-i-1} f(x) f(x+d) \dd x 
\end{equation}
\end{lemma}

\begin{proof}
This is just given by setting the integrand as $g_{i, i+1}(x, x+d)$.
We briefly note that the integral can only be evaluated from 0 to $1-d$ in this case because of the $x+d$ terms.
\end{proof}

\begin{lemma}[The cumulative distribution function of $D_i$]

\begin{equation}
\label{eq:cdf}
P(D_i < \delta) = H_i(\delta) = 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x
\end{equation}
\end{lemma}

\begin{proof}
$$
\begin{aligned}
H_i(\delta) & = \int_x^{x+\delta} h_i(d) \dd d \\
& = \int_x^{x+\delta} \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} ((F(x))^{i-1} (1 - F(x+d))^{n-i-1} f(x) f(x+d) \dd x \dd d \\
& = \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} f(x) \int_x^{x+\delta} (1 - F(x+d))^{n-i-1} f(x+d) \dd d \dd x \\
& = \int_0^{1} \frac{n!}{(i-1)! (n-i-1)!} (F(x))^{i-1} f(x) \int_{F(x)}^{F(x+\delta)} (1 - u)^{n-i-1} \dd u \dd x \\
& = \int_0^{1} \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} f(x) ((1 - F(x))^{n-i} - (1 - F(x + \delta))^{n-i}) \dd x \\ 
& = \int_0^1 g_i(x) \dd x - \int_0^1 \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x \\
& = 1 - \int_0^1 \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x \\
\end{aligned}
$$

Because of the $x + \delta$ term, we can't actually evaluate this integral all the way up to $1$, and so we are left with 
$$
\begin{aligned}
& = 1 - \int_0^{1 - \delta} \frac{n!}{(i-1)! (n-i)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x.
\end{aligned}
$$
\end{proof}

Thus if we are able to show that Eq. (\ref{eq:cdf}) converges to 1, we are done.

### Uniform case

\begin{lemma}[Differences between order statistics of a uniform distribution]
If $X_1, ..., X_n \iid \Uniform(0, 1)$, then each $D_i \sim \Betadist(1, n)$.
\end{lemma}

\begin{proof}
We begin with Eq. (\ref{eq:pdf}), plugging in $f(x) = 1$ and $F(x) = x$:

$$
h_i(d) = \int_0^{1-d} \frac{n!}{(i-1)! (n-i-1)!} x^{i-1} (1-x-d)^{n-i-1} \dd x
$$

Then we proceed with integration by parts, setting 
$u = x^{i-1} \implies du = (i-1) x^{i-2}$ and 
$dv = (1-x-d)^{n-i-1} dx \implies v = -\frac{1}{n-i} (1-x-d)^{n-i-1}$. 
Note that $u v |_0^{1-d} = 0$ in this case. This yields

$$
= \frac{n!}{(i-1)! (n-i-1)!} \int \frac{i-1}{n-i} x^{i-2} (1-x-d)^{n-i} \dd x
$$

Then applying integration by parts again until the $x^p$ term disappears, we get:

$$
\begin{aligned}
& = \frac{n!}{(i-1)! (n-i-1)!} \frac{(i-1)!}{(n-i) \cdots (n-2)} \int_0^{1-d} (1-x-d)^{n-2} \dd x \\
& = -\frac{n (n-1)}{n-1} (1-x-d)^{n-1} \Big|_0^{1-d} \\
& = n (1 - d)^{n-1}
\end{aligned}
$$

This the density function for $\Betadist(1, n)$, completing the proof.
\end{proof}

\begin{theorem}
Let $X_1, ..., X_n \iid \Uniform(0, 1)$. 
Then for any $\delta > 0$, $P(\max_i X_{(i+1)} - X_{(i)} < \delta) \to 1$ as 
$n \to \infty$.
\end{theorem}

\begin{proof}[Proof (sketch)]
Since $X_{(i+1)} - X_{(i)} = D_i \sim \Betadist(1, n)$, 
$P(X_{(i+1)} - X_{(i)} < \delta) = 1 - (1 - \delta)^n $. This yields

$$
\begin{aligned}
P(\max_i D_i < \delta) & \geq (P(D_i < \delta))^{n-1} \\
& = (1 - (1 - \delta)^n)^{n-1} \\
& \approx e^{-n \exp(-n \delta)}.
\end{aligned}
$$

In the limit $n \to \infty$, this goes to 1.
\end{proof}

### General case

\begin{theorem}
Let $X_1, ..., X_n \iid F$ with support $[0, 1]$, and suppose $f(x)$ is continuous and $f(x) \geq a > 0$ everywhere on the support. 
Let $D_i = X_{(i+1)} - X_{(i)}$. 
Then for any $\epsilon > 0$, there exists $N \in \mathbb{N}$ such that $P(\max_i D_i < \delta) > 1 - \epsilon$ when $n > N$.
\end{theorem}

\begin{proof}[Proof (sketch)]
We start with Eq. (\ref{eq:cdf}):

$$P(D_i \leq \delta) = 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x.$$

Making the approximation $F(x+\delta) \approx F(x) + \delta f(x)$ 
and bounding $f(x) \geq a$, we get:

$$P(D_i \leq \delta) \geq 1 - \int_0^{1-\delta} \frac{n!}{(n-i)! (i-1)!} (F(x))^{i-1} (1 - F(x) - a \delta)^{n-i} f(x) \dd x.$$

Then making the substitution $u = F(x) \implies du = f(x) \dd x$, we obtain

$$1 - \int_0^{F(1-\delta)} \frac{n!}{(n-i)! (i-1)!} u^{i-1} (1 - u - a \delta)^{n-i} \dd u$$

Evaluating the integral yields

$$P(D_i < \delta) = 1 - (1 - a \delta)^n + (1 - F(1-\delta) - a \delta)^n.$$

Then as before,

$$
\begin{aligned}
P(\max_i D_i < \delta) & \geq (P(D_i < \delta))^{n-1} \\
& \approx (1 - (1 - a \delta)^n + (1 - F(1-\delta) - a \delta)^n)^{n-1} \\
& \to 1
\end{aligned}
$$

\end{proof}

If $F$ is absolutely continuous, then it can be shown that

$$P(D_i \leq \delta) = 1 - \int_0^{1-\delta} \frac{n!}{(n-i+1)! (i-1)!} (F(x))^{i-1} (1 - F(x + \delta))^{n-i} f(x) \dd x$$

Making the approximation $F(x+\delta) \approx F(x) + \delta f(x)$ 
and bounding $f(x) \geq a > 0$, we get:

$$P(D_i \leq \delta) \geq 1 - \int_0^{1-\delta} \frac{n!}{(n-i+1)! (i-2)!} (F(x))^{i-2} (1 - F(x) - a \delta)^{n-i} f(x) dx$$

Then making the substitution $u = F(x) \implies du = f(x) dx$, we get

$$1 - \int_0^{F(1-\delta)} \frac{n!}{(n-i+1)! (i-2)!} u^{i-2} (1 - u - a \delta)^{n-i} du$$

Which becomes $1 - C P(n) \delta^{n}$ for some $C > 0$ and $P(n)$ is a polynomial of $n$, giving us the desired result $D_i \stackrel{p}{\to} 0$.

# Algorithms

TBD

# Computational Results

TBD