### 2021-10-20 curve-fitting-methods/curve-fitting-scratch.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
  11pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  colorlinks=true,
  linkcolor=Maroon,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=blue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{setspace}
\usepackage{float}
\usepackage{mathtools}
\usepackage{natbib}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\setcitestyle{numbers,square,comma}
\usepackage{verbatim}
\usepackage{amsthm}
\usepackage{comment}
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{plainnat}

\title{Semi-Parametric Manifold Clustering}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{Tr}}
\newcommand{\blockdiag}{\mathrm{blockdiag}}
\newcommand{\indep}{\stackrel{\mathrm{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
\newcommand{\Bernoulli}{\mathrm{Bernoulli}}
\newcommand{\Betadist}{\mathrm{Beta}}
\newcommand{\Uniform}{\mathrm{Uniform}}
\newcommand{\BG}{\mathrm{BernoulliGraph}}
\newcommand{\Categorical}{\mathrm{Categorical}}
\newcommand{\Multinomial}{\mathrm{Multinomial}}
\newcommand{\RDPG}{\mathrm{RDPG}}
\newcommand{\GRDPG}{\mathrm{GRDPG}}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\theoremstyle{example}
\newtheorem{example}{Example}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\as}{\stackrel{\mathrm{a.s.}}{\to}}

\hypertarget{estimating-polynomial-curves}{%
\section{Estimating Polynomial
Curves}\label{estimating-polynomial-curves}}

\hypertarget{problem-setup}{%
\subsection{Problem Setup}\label{problem-setup}}

Let:

\begin{itemize}
\tightlist
\item
  \(T_1, ..., T_n \stackrel{\mathrm{iid}}{\sim}F\) with support
  \([0, 1]\).
\item
  \(g(\cdot, \theta) : [0, 1] \mapsto \mathcal{X} \subset \mathbb{R}^d\).
\item
  \(X_1, ..., X_n = g(T_1), ..., g(T_n)\)
\end{itemize}

Assuming some parametric form of \(g\) with parameters \(\theta\), we
want to find \(\hat{\theta}\), some ``reasonable'' estimate for
\(\theta\). We observe \(X_i\) but not \(T_i\).

For now, we limit \(d = 2\) and \(g\) to quadratic functions.

\begin{example}

Let $g(t) = (t^2, 2 t (1 - t) ) = (t^2, 2 t - 2 t^2)$. 
(This is the first two dimensions of the Hardy-Weinberg curve). 
Then $\theta = (0, 0, 1, 0, 2, -2)$.

\begin{center}\includegraphics{curve-fitting-scratch_files/figure-latex/unnamed-chunk-3-1} \end{center}

\end{example}

If we observe the \(T_i\)'s, then we can use a standard polynomial
regression method to obtain \(\hat{\theta}\). Since we do not observe
them, the proposed iterative method is as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Initialize \(\hat{\theta}^{(0)}\) (e.g., by drawing from a probability
  distribution).
\item
  Estimate each \(\hat{t}_i^{(s)}\) by minimizing
  \(L(t_i, \hat{\theta}^{(s)} | x_i) = L_i = \|x_i - g(t_i | \hat{\theta}^{(s)})\|^2\).
\item
  Compute each
  \(\hat{x}_i^{(s)} = g(\hat{t}_i^{(s)} | \hat{\theta}^{(s)})\)
\item
  Estimate \(\hat{\theta}^{(s+1)}\) by minimizing
  \(L(\{\hat{t}_i^{(s)}\}, \theta | X) = \sum_i \|x_i - g(\hat{t}_i^{(s)} | \theta)\|^2\).
\item
  Repeat steps 2-4 until convergence.
\end{enumerate}

If we restrict \(g\) to be polynomials, then steps (2) and (4) have
closed-form solutions.

\begin{example}

Write $g(t | \theta) = (g_1(t | \theta_1), ..., g_d(t | \theta_d))$ where $g_r(t | \theta_r)$ is the component of $g$ in the $r^{th}$ dimension and $\theta_r$ is the vector of parameters for the $r^{th}$ dimension. If $g_r$ are polynomials of degree $p$, then each $\theta_r$ contains up to $p + 1$ entries.

Given the observed points $x_1, ..., x_n \in \mathbb{R}^d$ and their corresponding index points $t_1, ..., t_n \in \mathbb{R}$, we can find each $\hat{\theta}_r$ individually by $\hat{\theta}_r = A^{-1} b$ where $b \in \mathbb{R}^{p+1}$ and $b_k = \sum_i x_i t_i^k$ and $A \in \mathbb{R}^{(p+1) \times (p+1)}$ and $A_{kl} = \sum_i t^{(k-1) (l-1)}$.

On the other hand, if we have parameters $\theta$ but not the index points $t_i$, we can minimize each $t_i$ individually by finding the roots of a $p+1$ polynomial with coefficients that depend on $x_1, ..., x_n$ and $\theta$.

In the following plot, we drew $n = 200$ points from the 2D H-W curve with $T_1, ..., T_n \stackrel{\mathrm{iid}}{\sim}Uniform(0, 1)$. 
The red line is the curve that was fit using the above method.

\end{example}

\begin{center}\includegraphics{curve-fitting-scratch_files/figure-latex/unnamed-chunk-4-1} \end{center}

One problem with this method is the parameterization of the curve is not
unique.

\begin{example}

In the next example, we draw $A \sim \mathrm{RDPG}(X)$ using the same H-W curve and sample size as above and estimate the true latent positions (up to rotation). 
In this example, we force the intercept terms to be zero.

\end{example}

\begin{center}\includegraphics{curve-fitting-scratch_files/figure-latex/unnamed-chunk-6-1} \end{center}

Next, suppose we have K curves parameterized by \(g^{(k)}\), with points
drawn along these curves. Then one possible clustering technique is as
follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Assign an initial clustering (e.g., via spectral clustering).
\item
  Estimate the curve for each cluster.
\item
  Reassign the clusters by proximity to each cluster.
\item
  Repeat 2 and 3 until convergence.
\end{enumerate}

\begin{example}

We again limit these to be quadratic functions in $\mathbb{R}^2$. 
Here, $K = 2$ and $n_1 = n_2 = 256$.

\end{example}

\begin{center}\includegraphics{curve-fitting-scratch_files/figure-latex/unnamed-chunk-8-1} \end{center}

\begin{example}

Finally, we draw $A \sim \mathrm{RDPG}(X)$ from the above example and apply the clustering and curve fitting to the ASE of $A$.

\end{example}

\begin{verbatim}
   zhat.manifold
z     1   2
  1 187  69
  2  87 169
\end{verbatim}

\begin{center}\includegraphics{curve-fitting-scratch_files/figure-latex/unnamed-chunk-9-1} \end{center}

  \bibliography{misc.bib}

\end{document}