### 2021-11-10 curve-fitting-methods/curve-fitting-scratch.Rmd
updated notes

---
title: |
  | Semi-Parametric Manifold Clustering
output:
  pdf_document:
    keep_tex: true
    citation_package: natbib
fontsize: 11pt
# geometry: "left=1in,right=1in,top=1in,bottom=1in"
urlcolor: blue
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
bibliography: misc.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.pos = 'H', 
                      fig.align = 'center', 
                      fig.height = 3, 
                      fig.width = 3)
options(xtable.comment = FALSE, 
        xtable.table.placement = 'H')

library(ggplot2)
import::from(magrittr, `%>%`)
```

\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{Tr}}
\newcommand{\blockdiag}{\mathrm{blockdiag}}
\newcommand{\indep}{\stackrel{\mathrm{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
\newcommand{\Bernoulli}{\mathrm{Bernoulli}}
\newcommand{\Betadist}{\mathrm{Beta}}
\newcommand{\Uniform}{\mathrm{Uniform}}
\newcommand{\BG}{\mathrm{BernoulliGraph}}
\newcommand{\Categorical}{\mathrm{Categorical}}
\newcommand{\Multinomial}{\mathrm{Multinomial}}
\newcommand{\RDPG}{\mathrm{RDPG}}
\newcommand{\GRDPG}{\mathrm{GRDPG}}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\theoremstyle{example}
\newtheorem{example}{Example}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\as}{\stackrel{\mathrm{a.s.}}{\to}}

# Estimating Polynomial Curves

## Problem Setup

Let:

* $T_1, ..., T_n \iid F$ with support $[0, 1]$. 
* $g(\cdot, \theta) : [0, 1] \mapsto \mathcal{X} \subset \mathbb{R}^d$. 
* $X_1, ..., X_n = g(T_1), ..., g(T_n)$

Assuming some parametric form of $g$ with parameters $\theta$, we want to find $\hat{\theta}$, some "reasonable" estimate for $\theta$. 
We observe $X_i$ but not $T_i$.

For now, we limit $d = 2$ and $g$ to quadratic functions.

```{r}
source('~/dev/pabm-grdpg/functions.R')
source('~/dev/dissertation/curve-fitting-methods/functions.R')
```

\begin{example}

Let $g(t) = (t^2, 2 t (1 - t) ) = (0 + 0 t + t^2, 0 + 2 t - 2 t^2)$. 
(This is the first two dimensions of the Hardy-Weinberg curve). 
Then $\theta = (0, 0, 1, 0, 2, -2)$.

```{r}
t. <- seq(0, 1, .01)
X <- param.curve.quadr.2(t., 0, 0, 1, 0, 2, -2)

ggplot() + 
  geom_line(aes(x = X[, 1], y = X[, 2])) + 
  theme_bw() + 
  labs(x = expression(x[1]), y = expression(x[2])) + 
  coord_fixed()
```

\end{example}

If we observe the $T_i$'s, then we can use a standard polynomial regression method to obtain $\hat{\theta}$. Since we do not observe them, the proposed iterative method is as follows:

1. Initialize $\hat{\theta}^{(0)}$ (e.g., randomly). 
2. Estimate each $\hat{t}_i^{(s)}$ by minimizing $L(t_i, \hat{\theta}^{(s)} | x_i) = L_i = \|x_i - g(t_i | \hat{\theta}^{(s)})\|^2$.
3. Compute each $\hat{x}_i^{(s)} = g(\hat{t}_i^{(s)} | \hat{\theta}^{(s)})$
4. Estimate $\hat{\theta}^{(s+1)}$ by minimizing $L(\{\hat{t}_i^{(s)}\}, \theta | X) = \sum_i \|x_i - g(\hat{t}_i^{(s)} | \theta)\|^2$.
5. Repeat steps 2-4 until convergence.

If we restrict $g$ to be polynomials, then steps (2) and (4) have closed-form solutions. 
Alternatively, we can estimate $g$ using more general forms, e.g., splines, which may require approximation.

\begin{example}

Write $g(t | \theta) = (g_1(t | \theta_1), ..., g_d(t | \theta_d))$ where $g_r(t | \theta_r)$ is the component of $g$ in the $r^{th}$ dimension and $\theta_r$ is the vector of parameters for the $r^{th}$ dimension. If $g_r$ are polynomials of degree $p$, then each $\theta_r$ contains up to $p + 1$ entries.

Given the observed points $x_1, ..., x_n \in \mathbb{R}^d$ and their corresponding index points $t_1, ..., t_n \in \mathbb{R}$, we can find each $\hat{\theta}_r$ individually by $\hat{\theta}_r = A^{-1} b$ where $b \in \mathbb{R}^{p+1}$ and $b_k = \sum_i x_i t_i^k$ and $A \in \mathbb{R}^{(p+1) \times (p+1)}$ and $A_{kl} = \sum_i t^{(k-1) (l-1)}$.

On the other hand, if we have parameters $\theta$ but not the index points $t_i$, we can minimize each $t_i$ individually by finding the roots of a $p+1$ polynomial with coefficients that depend on $x_1, ..., x_n$ and $\theta$.

In the following plot, we drew $n = 200$ points from the 2D H-W curve with $T_1, ..., T_n \iid Uniform(0, 1)$. 
The red line is the curve that was fit using the above method.

\end{example}

```{r}
set.seed(314)
n <- 200
t. <- runif(n)
X <- param.curve.quadr.2(t., 0, 0, 1, 0, 2, -2)

out <- estimate.quadr.curve.2(X, maxit = 100)
thetahat <- out$theta
t.hat <- sort(out$t)
Xhat <- param.curve.quadr.2(t.hat, 
                            thetahat[1], thetahat[2], thetahat[3],
                            thetahat[4], thetahat[5], thetahat[6])

ggplot() + 
  geom_point(aes(x = X[, 1], y = X[, 2])) + 
  theme_bw() + 
  labs(x = expression(x[1]), y = expression(x[2])) + 
  coord_fixed() + 
  geom_line(aes(x = Xhat[, 1], y = Xhat[, 2]),
            colour = 'red')
```

Note: the parameterization of the curve is not unique.

# Estimation with Noise

\begin{example}

In the next example, we draw $A \sim \RDPG(X)$ using the same H-W curve and sample size as above and estimate the true latent positions (up to rotation).

\end{example}

```{r}
P <- X %*% t(X)
diag(P) <- 0
A <- draw.graph(P)
Y <- embedding(A, 2, 0)
```

```{r, cache = TRUE}
out1 <- estimate.quadr.curve.2(Y,
                              intercept = TRUE,
                              maxit = 200)
thetahat <- out1$theta
t.hat <- sort(out1$t)
Yhat <- param.curve.quadr.2(t.hat, 
                            thetahat[1], thetahat[2], thetahat[3],
                            thetahat[4], thetahat[5], thetahat[6])

ggplot() + 
  geom_point(aes(x = Y[, 1], y = Y[, 2])) + 
  theme_bw() + 
  labs(x = expression(x[1]), y = expression(x[2])) + 
  coord_fixed() + 
  geom_line(aes(x = Yhat[, 1], y = Yhat[, 2]),
            colour = 'red')
```

A modification to this that is possibly more robust is to use Bezier curves for $g$. 
This is the same functional form as the polynomial curves used before, but with orthogonal bases:

$$g(t | p) = \sum_{r = 0}^{R} p_r \binom{R}{r} (1 - t)^{R - r} t^r$$

where $R$ is the order of the Bezier curve, each $p_s$ is a vector of length $d$, and, as before, $g : [0, 1] \mapsto \mathbb{R}^d$. 
Thus, if we fit each $p_r$, then the procedure is the same as before.

The least squares estimate for $p \in \mathbb{R}^{R \times d}$ is

$$\hat{p} = (T^\top T)^{-1} T^\top X$$

where $T = \begin{bmatrix} t^{(1)} & \cdots & t^{(R)} \end{bmatrix}$ and each $t^{(r)}_i = \binom{R}{r} (1 - t_i)^{R - r} t_i^r$, and $X \in \mathbb{R}^{n \times d}$. 
The same procedure for estimating $t_1, ..., t_n$ can be applied here.

The parameterization for a given curve is not unique. 
In particular, the above procedure will not necessarily provide $t_1, ..., t_n \in [0, 1]$. 
One possible remedy for this is to, after estimating the $t_i$'s, normalize them to the unit interval. If we assume $t_1, ..., t_n \iid Uniform(a, b)$, then the UMVUE are

$$\hat{a} = \frac{n t_{(1)} - t_{(n)}}{n - 1},
\hat{b} = \frac{n t_{(n)} - t_{(1)}}{n - 1}$$

which yields the normalization transformation $t \leftarrow (t - \hat{a}) / (\hat{b} - \hat{a})$.

Alternatively, we can force the $t_i$'s to be approxiately uniform on the unit interval by the transformation $t \leftarrow \hat{F}(t)$, where $\hat{F}$ is the empirical CDF of $t_1, ..., t_n$.

For initialization, we can use a one-dimensional Isomap embedding to estimate the $t_i$'s and use that to estimate $\hat{p}$.
Experiments suggest that if the data are well-behaved (i.e., it looks like the curve we are trying to fit), this results in much faster convergence.

# Clustering

Next, suppose we have K curves parameterized by $g^{(k)}$, with points drawn along these curves. Then one possible clustering technique is as follows:

1. Assign an initial clustering (e.g., via spectral clustering).
2. Estimate the curve for each cluster (using the same curve-fitting procedure as before).
3. Reassign the clusters by proximity to each curve.
4. Repeat 2 and 3 until convergence.

\begin{example}

```{r}
f1 <- function(t) {
  x1 <- t ^ 2
  x2 <- 2 * t - 2 * t ^ 2
  return(cbind(x1, x2))
}

f2 <- function(t) {
  x1 <- 2 * t - 2 * t ^ 2
  x2 <- 1 - 2 * t + t ^ 2
  return(cbind(x1, x2))
}
```

We have two intersecting curves, $g_1(t) = \begin{bmatrix} t^2 & 2 t (1 - t) \end{bmatrix}^\top$ and $g_2(t) = \begin{bmatrix} 2 t (1 - t) & (1 - t) ^ 2 \end{bmatrix}^\top$. $n_1 = n_2 = 256$ points are drawn uniformly from each.

```{r}
set.seed(314159265)
n1 <- 2 ** 8
n2 <- n1
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))

a <- 1
b <- 1
t1 <- rbeta(n1, a, b)
t2 <- rbeta(n2, a, b)
T <- rbind(t1, t2)

X1 <- f1(t1)
X2 <- f2(t2)
X <- rbind(X1, X2)

ggplot() + 
  geom_point(aes(x = X[, 1], y = X[, 2], colour = factor(z))) + 
  coord_fixed() + 
  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
  theme_bw()

P <- X %*% t(X)
diag(P) <- 0
A <- draw.graph(P)
Xhat <- embedding(A, 2, 0)
# if (mean(Xhat[, 1]) < 0) Xhat[, 1] <- -Xhat[, 1]
```

We draw $A \sim \RDPG(X)$ and obtain the following ASE:

```{r, fig.height = 3}
ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], colour = factor(z))) + 
  coord_fixed() + 
  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
  theme_bw()
```

Fitting two quadratic Bezier curves to these data yields a community detection error rate of 10\%. 
In the following plot, the points are labeled according to their estimated labels.

```{r cache = TRUE, eval = TRUE}
set.seed(314159)
manifold.out <- manifold.clustering.quadr.2(A, 2, 
                                            # initialization = z,
                                            # initialization = 'spectral',
                                            initialization = 'random',
                                            method = 'bezier')

zhat.manifold <- manifold.out$z

t. <- seq(-1, 1, .01)
theta1 <- manifold.out$theta[[1]]
theta2 <- manifold.out$theta[[2]]
Y1 <- param.curve.quadr.2(t., 
                          theta1[1], theta1[2], theta1[3], 
                          theta1[4], theta1[5], theta1[6])
Y2 <- param.curve.quadr.2(t.,
                          theta2[1], theta2[2], theta2[3],
                          theta2[4], theta2[5], theta2[6])
Y <- rbind(Y1, Y2)
z. <- c(rep(1, length(t.)),
        rep(2, length(t.)))

ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], colour = factor(zhat.manifold))) + 
  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
  theme_bw() + 
  geom_path(aes(x = Y[, 1], y = Y[, 2],
                group = factor(z.))) +
  xlim(min(Xhat[, 1]), max(Xhat[, 1])) + ylim(min(Xhat[, 2]), max(Xhat[, 2])) + 
  coord_fixed()
```

\end{example}

## Model-Based Clustering

The CLT property of the ASE may allow us to approximate $W x - \hat{x} \sim \mathcal{N}(f_k(t), \Sigma_k)$.
One possibly reasonable assumption for $\Sigma_k$ is $\Sigma_k = \Sigma_k(t) =  \diag(\sigma_{k1}^2(t), ..., \sigma_{kd}^2(t))$. A plot of the squared errors of cluster 1 in the previous example reveals the following:

```{r, fig.height = 2, fig.width = 8}
theta1 <- manifold.out$theta[[1]]
theta2 <- manifold.out$theta[[2]]
t. <- manifold.out$t
X1 <- param.curve.quadr.2(t.[[1]], 
                          theta1[1], theta1[2], theta1[3],
                          theta1[4], theta1[5], theta1[6])
X2 <- param.curve.quadr.2(t.[[2]], 
                          theta2[1], theta2[2], theta2[3],
                          theta2[4], theta2[5], theta2[6])
# plot(Xhat[zhat.manifold == 1, 1] - X1[, 1])
# plot(Xhat[zhat.manifold == 1, 2] - X1[, 2])

p1 <- ggplot() + 
  geom_point(aes(x = t.[[1]], y = (Xhat[zhat.manifold == 1, 1] - X1[, 1]) ^ 2)) + 
  stat_smooth(aes(x = t.[[1]], y = (Xhat[zhat.manifold == 1, 1] - X1[, 1]) ^ 2)) + 
  theme_bw() + 
  labs(x = 't', y = expression((x[1] - hat(x)[1])^2))

p2 <- ggplot() + 
  geom_point(aes(x = t.[[1]], y = (Xhat[zhat.manifold == 1, 2] - X1[, 2]) ^ 2)) + 
  stat_smooth(aes(x = t.[[1]], y = (Xhat[zhat.manifold == 1, 2] - X1[, 2]) ^ 2)) + 
  theme_bw() + 
  labs(x = 't', y = expression((x[2] - hat(x)[2])^2))

gridExtra::grid.arrange(p1, p2, nrow = 1)
```

Fitting some curve to each $\sigma_{kr}^2(t)$ allows us to compute estimated probabilities of each point belonging to each cluster/manifold: $\hat{q}_{ik} = \mathcal{N}(x_i \mid g_k(\hat{t}_i), \Sigma_k(\hat{t}_i))$. 
Then we can use weighted least squares to estimate the parameters of $g$:

$$\hat{p}_k = (T^\top Q_k T)^{-1} T^\top Q_k X$$

where $Q_k = \diag(q_{1k}, ..., q_{nk})$.