### 2022-01-12 curve-fitting-methods/curve-fitting-scratch.Rmd
added example

---
title: |
  | Semi-Parametric Manifold Clustering
output:
  pdf_document:
    keep_tex: true
    citation_package: natbib
fontsize: 11pt
# geometry: "left=1in,right=1in,top=1in,bottom=1in"
urlcolor: blue
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
bibliography: misc.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.pos = 'H', 
                      fig.align = 'center', 
                      fig.height = 3, 
                      fig.width = 3)
options(xtable.comment = FALSE, 
        xtable.table.placement = 'H')

library(ggplot2)
import::from(magrittr, `%>%`, `%<>%`)
```

\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{Tr}}
\newcommand{\blockdiag}{\mathrm{blockdiag}}
\newcommand{\indep}{\stackrel{\mathrm{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
\newcommand{\Bernoulli}{\mathrm{Bernoulli}}
\newcommand{\Betadist}{\mathrm{Beta}}
\newcommand{\Uniform}{\mathrm{Uniform}}
\newcommand{\BG}{\mathrm{BernoulliGraph}}
\newcommand{\Categorical}{\mathrm{Categorical}}
\newcommand{\Multinomial}{\mathrm{Multinomial}}
\newcommand{\RDPG}{\mathrm{RDPG}}
\newcommand{\GRDPG}{\mathrm{GRDPG}}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\theoremstyle{example}
\newtheorem{example}{Example}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\as}{\stackrel{\mathrm{a.s.}}{\to}}
\newcommand{\ind}{\stackrel{\mathrm{ind}}{\sim}}

# Estimating Polynomial Curves

## Problem Setup

Let:

* $T_1, ..., T_n \iid F$ with support $[0, 1]$. 
* $g(\cdot, \theta) : [0, 1] \mapsto \mathcal{X} \subset \mathbb{R}^d$. 
* $X_1, ..., X_n = g(T_1), ..., g(T_n)$

Assuming some parametric form of $g$ with parameters $\theta$, we want to find $\hat{\theta}$, some "reasonable" estimate for $\theta$. 
We observe $X_i$ but not $T_i$.

For now, we limit $d = 2$ and $g$ to quadratic functions.

```{r}
source('~/dev/pabm-grdpg/functions.R')
source('~/dev/dissertation/curve-fitting-methods/functions.R')
source('https://mtrosset.pages.iu.edu/Courses/675/graph.r')
```

\begin{example}

Let $g(t) = (t^2, 2 t (1 - t) ) = (0 + 0 t + t^2, 0 + 2 t - 2 t^2)$. 
(This is the first two dimensions of the Hardy-Weinberg curve). 
Then $\theta = (0, 0, 1, 0, 2, -2)$.

```{r}
t. <- seq(0, 1, .01)
X <- param.curve.quadr.2(t., 0, 0, 1, 0, 2, -2)

ggplot() + 
  geom_line(aes(x = X[, 1], y = X[, 2])) + 
  theme_bw() + 
  labs(x = expression(x[1]), y = expression(x[2])) + 
  coord_fixed()
```

\end{example}

If we observe the $T_i$'s, then we can use a standard polynomial regression method to obtain $\hat{\theta}$. Since we do not observe them, the proposed iterative method is as follows:

1. Initialize $\hat{\theta}^{(0)}$ (e.g., randomly). 
2. Estimate each $\hat{t}_i^{(s)}$ by minimizing $L(t_i, \hat{\theta}^{(s)} | x_i) = L_i = \|x_i - g(t_i | \hat{\theta}^{(s)})\|^2$.
3. Compute each $\hat{x}_i^{(s)} = g(\hat{t}_i^{(s)} | \hat{\theta}^{(s)})$
4. Estimate $\hat{\theta}^{(s+1)}$ by minimizing $L(\{\hat{t}_i^{(s)}\}, \theta | X) = \sum_i \|x_i - g(\hat{t}_i^{(s)} | \theta)\|^2$.
5. Repeat steps 2-4 until convergence.

If we restrict $g$ to be polynomials, then steps (2) and (4) have closed-form solutions. 
Alternatively, we can estimate $g$ using more general forms, e.g., splines, which may require approximation.

\begin{example}

Write $g(t | \theta) = (g_1(t | \theta_1), ..., g_d(t | \theta_d))$ where $g_r(t | \theta_r)$ is the component of $g$ in the $r^{th}$ dimension and $\theta_r$ is the vector of parameters for the $r^{th}$ dimension. If $g_r$ are polynomials of degree $p$, then each $\theta_r$ contains up to $p + 1$ entries.

Given the observed points $x_1, ..., x_n \in \mathbb{R}^d$ and their corresponding index points $t_1, ..., t_n \in \mathbb{R}$, we can find each $\hat{\theta}_r$ individually by $\hat{\theta}_r = A^{-1} b$ where $b \in \mathbb{R}^{p+1}$ and $b_k = \sum_i x_i t_i^k$ and $A \in \mathbb{R}^{(p+1) \times (p+1)}$ and $A_{kl} = \sum_i t^{(k-1) (l-1)}$.

On the other hand, if we have parameters $\theta$ but not the index points $t_i$, we can minimize each $t_i$ individually by finding the roots of a $p+1$ polynomial with coefficients that depend on $x_1, ..., x_n$ and $\theta$.

In the following plot, we drew $n = 200$ points from the 2D H-W curve with $T_1, ..., T_n \iid Uniform(0, 1)$. 
The red line is the curve that was fit using the above method.

\end{example}

```{r}
set.seed(314)
n <- 200
t. <- runif(n)
X <- param.curve.quadr.2(t., 0, 0, 1, 0, 2, -2)

out <- estimate.quadr.curve.2(X, maxit = 100)
thetahat <- out$theta
t.hat <- sort(out$t)
Xhat <- param.curve.quadr.2(t.hat, 
                            thetahat[1], thetahat[2], thetahat[3],
                            thetahat[4], thetahat[5], thetahat[6])

ggplot() + 
  geom_point(aes(x = X[, 1], y = X[, 2])) + 
  theme_bw() + 
  labs(x = expression(x[1]), y = expression(x[2])) + 
  coord_fixed() + 
  geom_line(aes(x = Xhat[, 1], y = Xhat[, 2]),
            colour = 'red')
```

Note: the parameterization of the curve is not unique.

# Estimation with Noise

\begin{example}

In the next example, we draw $A \sim \RDPG(X)$ using the same H-W curve and sample size as above and estimate the true latent positions (up to rotation).

\end{example}

```{r}
P <- X %*% t(X)
diag(P) <- 0
A <- draw.graph(P)
Y <- embedding(A, 2, 0)
```

```{r, cache = TRUE}
out1 <- estimate.quadr.curve.2(Y,
                              intercept = TRUE,
                              maxit = 200)
thetahat <- out1$theta
t.hat <- sort(out1$t)
Yhat <- param.curve.quadr.2(t.hat, 
                            thetahat[1], thetahat[2], thetahat[3],
                            thetahat[4], thetahat[5], thetahat[6])

ggplot() + 
  geom_point(aes(x = Y[, 1], y = Y[, 2])) + 
  theme_bw() + 
  labs(x = expression(x[1]), y = expression(x[2])) + 
  coord_fixed() + 
  geom_line(aes(x = Yhat[, 1], y = Yhat[, 2]),
            colour = 'red')
```

A modification to this that is possibly more robust is to use Bezier curves for $g$. 
This is the same functional form as the polynomial curves used before, but with orthogonal bases:

$$g(t | p) = \sum_{r = 0}^{R} p_r \binom{R}{r} (1 - t)^{R - r} t^r$$

where $R$ is the order of the Bezier curve, each $p_s$ is a vector of length $d$, and, as before, $g : [0, 1] \mapsto \mathbb{R}^d$. 
Thus, if we fit each $p_r$, then the procedure is the same as before.

The least squares estimate for $p \in \mathbb{R}^{R \times d}$ is

$$\hat{p} = (T^\top T)^{-1} T^\top X$$

where $T = \begin{bmatrix} t_{\cdot, 1} & \cdots & t_{\cdot, R} \end{bmatrix}$ and each $t_{ir} = \binom{R}{r} (1 - t_i)^{R - r} t_i^r$, and $X \in \mathbb{R}^{n \times d}$. 
The same procedure for estimating $t_1, ..., t_n$ can be applied here.

The parameterization for a given curve is not unique. 
In particular, the above procedure will not necessarily provide $t_1, ..., t_n \in [0, 1]$. 
One possible remedy for this is to, after estimating the $t_i$'s, normalize them to the unit interval. If we assume $t_1, ..., t_n \iid Uniform(a, b)$, then the UMVUE are

$$\hat{a} = \frac{n t_{(1)} - t_{(n)}}{n - 1},
\hat{b} = \frac{n t_{(n)} - t_{(1)}}{n - 1}$$

which yields the normalization transformation $t \leftarrow (t - \hat{a}) / (\hat{b} - \hat{a})$.

Alternatively, we can force the $t_i$'s to be approxiately uniform on the unit interval by the transformation $t \leftarrow \hat{F}(t)$, where $\hat{F}$ is the empirical CDF of $t_1, ..., t_n$.

For initialization, we can use a one-dimensional Isomap embedding to estimate the $t_i$'s and use that to estimate $\hat{p}$.
Experiments suggest that if the data are well-behaved (i.e., it looks like the curve we are trying to fit), this results in much faster convergence.

## Theoretical Framework

If we assume that the points are distributed as $X_i \ind \mathcal{N}(f(t_i | p), \Sigma(t_i | \phi))$, then we can write the (incomplete) log likelihood as:

$$\ell(p, \phi) = -\frac{1}{2} \sum_i \log |\Sigma(t_i | \phi)| - \frac{1}{2} \sum_i (x_i - f(t_i | p))^\top (\Sigma(t_i | \phi))^{-1} (x_i - f(t_i | p))$$

In the case $\Sigma(t_i | \phi) = \phi I$ and $f(t_i | p) = p^\top \tilde{t}_i$ where $\tilde{t}_i \in \mathbb{R}^{R+1}$ is the Bezier polynomial expansion of order $R$, then this becomes

$$\ell(p, \phi) = -\frac{n d}{2} \log \phi - \frac{1}{2 \phi} \|X - T p\|^2_F$$

Then given $t_1, ..., t_n$, the MLE of $p$ is again $(T^\top T)^{-1} T^\top X$, and the MLE of $\phi$ is just $\frac{1}{n d} \|X - T \hat{p}\|^2_F$.
And given $p$, we can estimate each $t_i$ in the same way as before to maximize $\ell$ (the variance is unnecessary here).

As noted before, the parameterization given by $T$ and $p$ is not unique. 
We can arbitrarily scale the $t_i$'s and $p$ to obtain the same value of $\ell$. 
Naively performing the proposed algorithm sometimes results in estimates for $p$ that diverge and estimates for $t_i$'s that converge to a single value. 
To remedy this, we can either scale the $t_i$'s as before, 
or we can scale $p$ by forcing $f$ to be an arclength parameterization.

\begin{theorem}
Let $x_1, ..., x_n \ind \mathcal{N}(p^\top \tilde{t}_i, \phi I_d)$,
where each $x_i \in \mathbb{R}^d$, $\tilde{t}_i \in \mathbb{R}^{R+1}$ such that $\tilde{t}_{i,r} = \binom{R}{r} (1 - t_i)^{R-r} t_i^r$ for $i = 1, ..., n$ and $r = 0, ..., R$, 
and $p \in \mathbb{R}^{(R+1) \times d}$.

Then each iteration of the following decreases the negative log likelihood:

\begin{enumerate}
  \item $p \leftarrow (T^\top T)^{-1} T^\top X$, 
  where $T = \begin{bmatrix} \tilde{t}_1 & \cdots & \tilde{t}_n \end{bmatrix}^\top$ 
  and $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$.
  \item $t_i \leftarrow \arg\min_t \|x_i - p^\top \tilde{t}_i\|^2$,
  which can be solved by finding the roots of a polynomial of degree $2 R - 1$.
\end{enumerate}
\end{theorem}

\begin{proof}[Proof (sketch)]

For $\Sigma(t_i | \phi) = \phi I$, the negative log likelihood is, 
up to some additive and multiplicative constants:

$$l(p, \{t_i\}) = \|X - T p\|^2_F = X^\top X - 2 X^\top T p + p^\top T^\top T p$$

To find the minimizer $\hat{p}$ given $t_1, ..., t_n$:

$\nabla_p l = -2 T^\top X + 2 T^\top T p = 0$  
$\implies \hat{p} = (T^\top T)^{-1} T^\top X$

To find the minimizer $\hat{t}_i$ given $p$:

We can rewrite the negative log likelihood as 
$l(p, \{t_i\}) = \sum_i \|x_i - p^\top \tilde{t}_i\|^2$.

Then each entry of $\nabla_{t} l$ depends only on $t_i$, so each $t_i$ can be optimized independently. 
Furthermore, each $\frac{\partial}{\partial t_i} \|x_i - p^\top \tilde{t}_i\|^2$ is a polynomial in $t_i$, so it can be minimized by simply finding the roots of the polynomial.

Thus this algorithm is a coordinate descent algorithm, which reduces the objective function with each iteration. 
\end{proof}

\begin{theorem}

Let $A^{(n)} \sim \RDPG(F(p, \theta), n)$, and let $X^{(n)}$ be the ASE of $A^{(n)}$. 
Let $\hat{p}, \{\hat{t}_i\}$ be the global minimizer of $l$ given $X^{(n)}$. 
Then $l(\hat{p}, \{\hat{t}_i\}) \stackrel{a.s.}{\to} 0$.

\end{theorem}

### Covariance Structures

While the CLT property of the ASE allows us to approximate $x_i \sim \mathcal{N} (W f(t_i | p), W \Sigma(t_i | \phi) W^\top)$, it is unclear what $\Sigma(t_i | \phi)$ looks like, and we cannot in general say $\Sigma(t_i | \phi) = \phi I_d$.

# Clustering

Next, suppose we have K curves parameterized by $g^{(k)}$, with points drawn along these curves. Then one possible clustering technique is as follows:

1. Assign an initial clustering (e.g., via spectral clustering).
2. Estimate the curve for each cluster (using the same curve-fitting procedure as before).
3. Reassign the clusters by proximity to each curve.
4. Repeat 2 and 3 until convergence.

\begin{example}

```{r}
f1 <- function(t) {
  x1 <- t ^ 2
  x2 <- 2 * t - 2 * t ^ 2
  return(cbind(x1, x2))
}

f2 <- function(t) {
  x1 <- 2 * t - 2 * t ^ 2
  x2 <- 1 - 2 * t + t ^ 2
  return(cbind(x1, x2))
}
```

We have two intersecting curves, $g_1(t) = \begin{bmatrix} t^2 & 2 t (1 - t) \end{bmatrix}^\top$ and $g_2(t) = \begin{bmatrix} 2 t (1 - t) & (1 - t) ^ 2 \end{bmatrix}^\top$. $n_1 = n_2 = 256$ points are drawn uniformly from each.

```{r}
set.seed(314159265)
n1 <- 2 ** 8
n2 <- n1
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))

a <- 1
b <- 1
t1 <- rbeta(n1, a, b)
t2 <- rbeta(n2, a, b)
T <- rbind(t1, t2)

X1 <- f1(t1)
X2 <- f2(t2)
X <- rbind(X1, X2)

ggplot() + 
  geom_point(aes(x = X[, 1], y = X[, 2], colour = factor(z))) + 
  coord_fixed() + 
  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
  theme_bw()

P <- X %*% t(X)
diag(P) <- 0
A <- draw.graph(P)
Xhat <- embedding(A, 2, 0)
# if (mean(Xhat[, 1]) < 0) Xhat[, 1] <- -Xhat[, 1]
```

We draw $A \sim \RDPG(X)$ and obtain the following ASE:

```{r, fig.height = 3}
ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], colour = factor(z))) + 
  coord_fixed() + 
  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
  theme_bw()
```

Fitting two quadratic Bezier curves to these data yields a community detection error rate of 10\%. 
In the following plot, the points are labeled according to their estimated labels.

```{r cache = TRUE, eval = TRUE}
set.seed(314159)
manifold.out <- manifold.clustering.quadr.2(A, 2, 
                                            # initialization = z,
                                            # initialization = 'spectral',
                                            initialization = 'random',
                                            method = 'bezier')

zhat.manifold <- manifold.out$z

t. <- seq(-1, 1, .01)
theta1 <- manifold.out$theta[[1]]
theta2 <- manifold.out$theta[[2]]
Y1 <- param.curve.quadr.2(t., 
                          theta1[1], theta1[2], theta1[3], 
                          theta1[4], theta1[5], theta1[6])
Y2 <- param.curve.quadr.2(t.,
                          theta2[1], theta2[2], theta2[3],
                          theta2[4], theta2[5], theta2[6])
Y <- rbind(Y1, Y2)
z. <- c(rep(1, length(t.)),
        rep(2, length(t.)))

ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], colour = factor(zhat.manifold))) + 
  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
  theme_bw() + 
  geom_path(aes(x = Y[, 1], y = Y[, 2],
                group = factor(z.))) +
  xlim(min(Xhat[, 1]), max(Xhat[, 1])) + ylim(min(Xhat[, 2]), max(Xhat[, 2])) + 
  coord_fixed()
```

\end{example}

The objective function that this method aims to minimize is:

$$\sum_{k=1}^K \sum_{i \in C_k} \|x_i - g_k(t_i | p_k)\|^2$$

Where $g_k : [0, 1] \mapsto \mathbb{R}^d$ with parameters $p_k$ is the curve for community $k$.

**TBD**: It can be shown that each iteration decreases the objective unless we are at a stationary point.

## Harry Potter Emnity Graph

\begin{example}
In this section, we analyze the Harry Potter emnity graph. 
Rubin-Delanchy et al. suggested applying the GRDPG ASE to this graph with $p = 1$ and $q = 1$, revealing a DCBM-like structure in the latent space. 
Passino and Heard instead fit Gaussian processes to the embedding. 
Here, we consider what happens when we apply the Bezier curve clustering method. 
\end{example}

```{r, fig.width = 8}
relations.df <- readr::read_csv('~/dev/potter-network/data/relations.csv') %>% 
  dplyr::mutate(source = source + 1, target = target + 1)
characters.df <- readr::read_csv('~/dev/potter-network/data/characters.csv') %>% 
  dplyr::mutate(id = id + 1)
n <- nrow(characters.df)

A <- matrix(0, nrow = n, ncol = n)
rownames(A) <- colnames(A) <- characters.df$name
for (i in seq(nrow(relations.df))) {
  if (relations.df$type[i] == '-') {
    A[relations.df$source[i], relations.df$target[i]] %<>% magrittr::add(1)
  } else {
    A[relations.df$source[i], relations.df$target[i]] %<>% magrittr::subtract(1)
  }
}
A[A < 0] <- 0
A <- A + t(A)
A <- sign(A)
diag(A) <- 0
deg.A <- colSums(A)

A <- A[deg.A > 0, deg.A > 0]

A.eigen <- eigen(A, symmetric = TRUE)
Xhat <- embedding(A, p = 1, q = 1)

# ggplot() + 
#   geom_point(aes(x = Xhat[, 1], y = Xhat[, 2])) + 
#   ggrepel::geom_text_repel(aes(x = Xhat[, 1], y = Xhat[, 2], 
#                                label = characters.df$name[deg.A > 0]))

z.init <- kmeans(sweep(Xhat, 1, apply(Xhat, 1, function(x) sqrt(sum(x ^ 2))), `/`), 2)$cluster
```

```{r, fig.width = 10, fig.height = 4, cache = TRUE}
manifold.out <- manifold.clustering.quadr.2(A, 2, p = 1, q = 1, 
                                            initialization = z.init, 
                                            eps.isomap = .7,
                                            method = 'bezier')

zhat.manifold <- manifold.out$z

t. <- seq(-1, 1, .01)
theta1 <- manifold.out$theta[[1]]
theta2 <- manifold.out$theta[[2]]
Y1 <- param.curve.quadr.2(t., 
                          theta1[1], theta1[2], theta1[3], 
                          theta1[4], theta1[5], theta1[6])
Y2 <- param.curve.quadr.2(t.,
                          theta2[1], theta2[2], theta2[3],
                          theta2[4], theta2[5], theta2[6])
Y <- rbind(Y1, Y2)
z. <- c(rep(1, length(t.)),
        rep(2, length(t.)))

ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], colour = factor(zhat.manifold))) + 
  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
  theme_bw() + 
  theme(legend.position = 'none') +
  geom_path(aes(x = Y[, 1], y = Y[, 2],
                group = factor(z.))) +
  ggrepel::geom_text_repel(aes(x = Xhat[, 1], y = Xhat[, 2],
                               label = characters.df$name[deg.A > 0])) + 
  xlim(min(Xhat[, 1]), max(Xhat[, 1])) + ylim(min(Xhat[, 2]), max(Xhat[, 2]))
```