### 2023-01-13 slides/slides.Rmd
finished adding references

---
title: Community Detection in the Setting of Generalized Random Dot Product Graphs
author: |
  | John Koo
  |
  | Department of Statistics
  | Indiana University
date: 'January 2023'
output: 
  beamer_presentation:
    fig_crop: no
    theme: 'default'
    colortheme: 'beaver'
    includes:
      in_header: page_headers.tex
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
- \setbeamertemplate{itemize items}[circle]
- \usepackage{caption}
- \captionsetup[figure]{font=scriptsize}
- \captionsetup[figure]{labelformat=empty}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE,
                      fig.align = 'center',
                      fig.lp = '')
options(xtable.comment = FALSE, 
        xtable.table.placement = 'H')
library(ggplot2)
library(ggraph)
import::from(magrittr, `%>%`)
import::from(foreach, `%dopar%`)
theme_set(theme_bw())

source('~/dev/pabm-grdpg/functions.R')
source('~/dev/manifold-block-models/functions.R')
set.seed(314159)

doMC::registerDoMC(parallel::detectCores())
```

## Outline

1. Random Graph Models
    * Block Models
    * Generalized Random Dot Product Graphs

2. PABMs are GRDPGs

3. Manifold Block Models

# Random Graph Models

## Problem Setup

\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{Tr}}
\newcommand{\blockdiag}{\mathrm{blockdiag}}
\newcommand{\indep}{\stackrel{\mathrm{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
\newcommand{\Bernoulli}{\mathrm{Bernoulli}}
\newcommand{\Betadist}{\mathrm{Beta}}
\newcommand{\BG}{\mathrm{BernoulliGraph}}
\newcommand{\Uniform}{\mathrm{Uniform}}
\newcommand{\Categorical}{\mathrm{Categorical}}
\newcommand{\PABM}{\mathrm{PABM}}
\newcommand{\RDPG}{\mathrm{RDPG}}
\newcommand{\GRDPG}{\mathrm{GRDPG}}
\newcommand{\Multinomial}{\mathrm{Multinomial}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\as}{\stackrel{\mathrm{a.s.}}{\to}}
\newcommand{\ER}{\text{Erd\"{o}s-R\'{e}nyi}}
\newcommand{\SBM}{\mathrm{SBM}}
\newcommand{\DCBM}{\mathrm{DCBM}}
\newcommand{\rank}{\mathrm{rank}}
\newcommand{\MBM}{\mathrm{MBM}}

```{r out.width = '50%', fig.height = 3, fig.width = 4, fig.cap = 'Friendship network of 81 faculty at a UK university (Nepusz et al., 2008). The vertices are labeled by school affiliation.'}
data('UKfaculty', package = 'igraphdata')
A <- as.matrix(igraph::as_adjacency_matrix(UKfaculty, type = 'both'))
A <- sign(A + t(A))
z <- factor(igraph::vertex_attr(UKfaculty)$Group)
n <- length(z)

ggraph(A, layout = 'stress') + 
  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
  geom_node_point(aes(colour = z), 
                  show.legend = FALSE) + 
  labs(colour = NULL) + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')
```
\begin{center}
How might we cluster the nodes of a network?
\end{center}

## Bernoulli Random Graphs

Let $G$ be an undirected and unweighted graph with $n$ vertices.

$G$ is described by adjacency matrix $A$ such that
$A_{ij} = \begin{cases} 
1 & \text{an edge connects vertices } i \text{ and } j \\
0 & \text{otherwise}
\end{cases}$

$A_{ji} = A_{ij}$ and $A_{ii} = 0$.

\vspace*{1\baselineskip}

$A \sim \BG(P)$ iff:

1. $P$ is a matrix of edge probabilities between pairs of vertices.
2. $A_{ij} \indep \Bernoulli(P_{ij})$ for each $i < j$.

## Block Models

Suppose each vertex $v_1, ..., v_n$ has labels $z_1, ..., z_n \in \{1, ..., K\}$,  
and each $P_{ij}$ depends on labels $z_i$ and $z_j$.  
Then $A \sim \BG(P)$ is a *block model*.

**Example 1**: Stochastic Block Model with $K=2$ communities.

::: columns

:::: column

$P_{ij} = \begin{cases} 
p & z_i = z_j = 1 \\
q & z_i = z_j = 2 \\
r & z_i \neq z_j
\end{cases}$

::::

:::: column

```{r, fig.height = 3, fig.width = 4, out.width = '100%'}
set.seed(123)

n1 <- 2 ** 5
n2 <- 2 ** 5
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))
p <- 1/2
q <- 1/4
r <- 1/8
P <- matrix(r, nrow = n, ncol = n)
P[seq(n1), seq(n1)] <- p
P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
A <- draw.graph(P)
qgraph::qgraph(A, vsize = 4, groups = factor(z), legend = FALSE)
```

::::

:::

## Block Models

Stochastic Block Model (Lorrain and White, 1971)

* $P_{ij} = \theta_{z_i z_j}$
* $K (K + 1) / 2$ parameters $\theta_{k \ell}$

Degree Corrected Block Model (Karrer and Newman, 2011)

* $P_{ij} = \theta_{z_i z_j} \omega_i \omega_j$
* $K (K + 1) / 2 + n$ parameters $\theta_{k \ell}$, $\omega_i$

Popularity Adjusted Block Model (Sengupta and Chen, 2018)

* $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$
* $K n$ parameters $\lambda_{ik}$

## Block Models

Dating network as an SBM: $P_{ij} = \theta_{z_i, z_j}$

* This model is disassortative; $\theta_{k \ell} > \theta_{kk}$ for $k \neq \ell$.
* Each individual has the same probability of matching with every other individual conditioned on gender.

```{r dating-sbm, fig.width = 4, , fig.height = 2, cache = TRUE}
set.seed(123)

n <- 2 ** 6
z <- sample(c(1, 2), n, replace = TRUE)
z <- sort(z)
n1 <- sum(z == 1)
n2 <- sum(z == 2)
z[z == 1] <- 'F'
z[z == 2] <- 'M'
theta11 <- .02
theta22 <- .02
theta12 <- .2
P <- matrix(theta12, nrow = n, ncol = n)
P[seq(n1), seq(n1)] <- theta11
P[seq(n1 + 1, n), seq(n1 + 1, n)] <- theta22
A <- draw.graph(P)
# qgraph::qgraph(A, vsize = 4, groups = factor(z), legend = FALSE)
ggraph(A, layout = 'auto') + 
  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
  geom_node_point(aes(colour = factor(z)), 
                  show.legend = TRUE) + 
  labs(colour = 'Gender') + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')
```

## Block Models

Dating network as a DCBM: $P_{ij} = \omega_i \omega_j \theta_{z_i, z_j}$

* Models each individual's popularity---some individuals form more connections than others.
* Still assumes that all members of each gender have the same odds of matching between two genders.

```{r dating-dcbm, fig.width = 4, , fig.height = 2, cache = TRUE}
set.seed(12345)

n <- 2 ** 6
z <- sample(c(1, 2), n, replace = TRUE)
z <- sort(z)
n1 <- sum(z == 1)
n2 <- sum(z == 2)
z[z == 1] <- 'F'
z[z == 2] <- 'M'
theta11 <- .03
theta22 <- .03
theta12 <- .3
omega <- runif(n, 1/3, 1)
Omega <- omega %*% t(omega)
P <- matrix(theta12, nrow = n, ncol = n)
P[seq(n1), seq(n1)] <- theta11
P[seq(n1 + 1, n), seq(n1 + 1, n)] <- theta22
P <- P * Omega
A <- draw.graph(P)
ggraph(A, layout = 'auto') + 
  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
  geom_node_point(aes(colour = factor(z)), 
                  show.legend = TRUE) + 
  labs(colour = 'Gender') + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')
```

## Block Models

Dating network as a PABM: $P_{ij} = \lambda_{i, z_j} \lambda_{j, z_i}$

* Models each user's sexual orientation as well as popularity. 
* Some individuals are more likely to connect with members of the same gender while others are more likely to connect with members of the opposite gender.

```{r dating-pabm, fig.width = 4, , fig.height = 2, cache = TRUE}
set.seed(12345)
n <- 2 ** 6
Az <- draw.pabm.beta.2(n, 1, 6, 2, 2, .5)
A <- Az$A
z <- Az$z
z[z == 1] <- 'F'
z[z == 2] <- 'M'

ggraph(A, layout = 'auto') + 
  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
  geom_node_point(aes(colour = factor(z)), 
                  show.legend = TRUE) + 
  labs(colour = 'Gender') + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')
```

## Hierarchy of Block Models

PABM $\to$ DCBM: $\lambda_{ik} = \sqrt{\theta_{z_i k}} \omega_i$

DCBM $\to$ SBM: $\omega_i = 1$

SBM $\to$ Erdos-Renyi: $\theta_{k \ell} = \theta$

\begin{center}
```{r, out.width = '125px'}
knitr::include_graphics('../proposal/noroozi-pensky-hierarchy.png')
```

\tiny{Noroozi and Pensky, 2021}
\end{center}

## Estimation for Block Models

Log-likelihood function for the SBM:

$$\ell(z, \theta; A) = \sum_{i<j} \sum_{k, \ell} z_{ik} z_{j \ell} \Big(A_{ij} \log \theta_{k \ell} + (1 - A_{ij}) \log(1 - \theta_{k \ell}) \Big)$$

* Direct maximization is NP-hard.

* Expectation-Maximization requires an independence relaxation.

## Estimation for Block Models

\begin{algorithm}[H]
  \label{algo:sbm-em}
  \scriptsize
  \DontPrintSemicolon
  \SetAlgoLined
  \KwData{Adjacency matrix $A$, number of communities $K$}
  \KwResult{Estimated community label probabilities $\{\pi_{ik}\}$ for which each $\pi_{ik} = P(z_i = k \mid A)$, estimated community edge probabilities $\{\hat{\theta}_{k \ell}\}_K$}
  \caption{Approximate EM algorithm for the SBM}
  Initialize $\{\pi_{ik}\}$, $\{\theta_{k \ell}\}$.\;
  \While{$\|\nabla \ell\| > \epsilon$} {
    \For {$i = 1, ..., n$} {
      \For {$k = 1, ..., K$} {
      E-step: $\pi_{ik} \propto \exp \bigg( \sum_{j \neq i} \sum_{\ell} \pi_{j \ell} (A_{ij} \log \hat{\theta}_{k \ell} + (1 - A_{ij}) \log (1 - \hat{\theta}_{k \ell})) \bigg)$.\;
      }
    }
    \For {$k = 1, ..., K$} {
      \For {$\ell = 1, ..., K$} {
        M-step: $\hat{\theta}_{k \ell} = \frac{\sum_{i < j} A_{ij} \pi_{ik} \pi_{j \ell}}{\sum_{i < j} \pi_{ik} \pi_{j \ell}}$.\;
      }
    }
  }
\end{algorithm}

## Generalized Random Dot Product Graphs

**Def** Generalized Random Dot Product Graph  
(Rubin-Delanchy, Cape, Tang, Priebe, 2022)

Let $I_{p,q} = \blockdiag(I_p, -I_q)$ and suppose that $x_1, \ldots, x_n \in \mathbb{R}^{p+q}$ are such that $x_i^\top I_{p,q} x_j \in [0,1]$.

Then $A \sim \GRDPG_{p, q}(X)$ iff $A \sim \BG(X I_{p,q} X^\top)$, where $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$.

```{r rdpg-example, fig.length = 6, fig.height = 2, out.width = '100%', cache = TRUE, fig.cap = 'Latent vectors (left) and GRDPG (right).'}
set.seed(123)

n <- 200
t. <- runif(n)
x <- t. ^ 2
y <- 2 * t. * (1 - t.)

latent.vec.plot <- ggplot() + 
  geom_point(aes(x = x, y = y, colour = t.)) + 
  coord_fixed() + 
  theme_bw() + 
  labs(x = NULL, y = NULL) + 
  theme(legend.position = 'none') + 
  viridis::scale_colour_viridis()

X <- cbind(x, y)
P <- X %*% t(X)
A <- draw.graph(P)

rdpg.plot <- ggraph(A, layout = 'auto') + 
  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
  geom_node_point(aes(colour = t.), 
                  show.legend = FALSE) + 
  theme_void() + 
  viridis::scale_colour_viridis()

gridExtra::grid.arrange(latent.vec.plot, rdpg.plot, nrow = 1)
```

## Generalized Random Dot Product Graphs

Adjacency Spectral Embedding (Sussman et al., 2012) estimates $x_1, ..., x_n \in \mathbb{R}^{p+q}$ from $A$:

1. Let $\hat{\Lambda}$ be the diagonal matrix that contains the absolute values of the $p$ most positive and the $q$ most negative eigenvalues.
2. Let $\hat{V}$ be the matrix whose columns are the corresponding eigenvectors.
3. Compute $\hat{X} = \hat{V} \hat{\Lambda}^{1/2}$.

**Theorem**: $\max\limits_i \|\hat{X}_i - Q_n X_i \| = O_P \Big( \frac{(\log n)^c}{n^{1/2}} \Big)$ 
as $n \to \infty$

```{r ase-example, fig.length = 8, fig.height = 2, out.width = '100%', cache = TRUE, fig.cap = 'Latent vectors (left), GRDPG (center), and ASE (right).'}
set.seed(123)

n <- 200
t. <- runif(n)
x <- t. ^ 2
y <- 2 * t. * (1 - t.)

latent.vec.plot <- ggplot() + 
  geom_point(aes(x = x, y = y, colour = t.)) + 
  coord_fixed() + 
  theme_bw() + 
  labs(x = NULL, y = NULL) + 
  theme(legend.position = 'none') + 
  viridis::scale_colour_viridis()

X <- cbind(x, y)
P <- X %*% t(X)
A <- draw.graph(P)

rdpg.plot <- ggraph(A, layout = 'auto') + 
  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
  geom_node_point(aes(colour = t.), 
                  show.legend = FALSE) + 
  theme_void() + 
  viridis::scale_colour_viridis()

Y <- embedding(P, 2, 0)
Xhat <- embedding(A, 2, 0)
Xhat[, 1] <- -Xhat[, 1]
ase.plot <- ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], colour = t.)) + 
  geom_line(aes(x = Y[, 1], y = Y[, 2]), colour = 'red') + 
  coord_fixed() + 
  theme_bw() + 
  labs(x = NULL, y = NULL) + 
  theme(legend.position = 'none') + 
  viridis::scale_colour_viridis()

gridExtra::grid.arrange(latent.vec.plot, rdpg.plot, ase.plot, nrow = 1)
```

## SBMs are GRDPGs

All Bernoulli Graphs are GRDPGs.

**Example 1** (cont'd): SBM with $K = 2$.

::: columns

:::: column

$$P_{ij} = \begin{cases} 
p & z_i = z_j = 1 \\
q & z_i = z_j = 2 \\
r & z_i \neq z_j
\end{cases}$$

```{r fig.height = 3, fig.width = 6, out.width = '100%'}
set.seed(123)

n1 <- 2 ** 5
n2 <- 2 ** 5
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))
p <- 1/2
q <- 1/4
r <- 1/8
P <- matrix(r, nrow = n, ncol = n)
P[seq(n1), seq(n1)] <- p
P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
A <- draw.graph(P)
qgraph::qgraph(A, vsize = 4, groups = factor(z), legend = FALSE)
```

::::

:::: column

$$P = 
\begin{bmatrix} 
P^{(11)} & P^{(12)} \\
P^{(21)} & P^{(22)}
\end{bmatrix} =
X I_{2,0} X^\top$$

$$X = \begin{bmatrix} 
\sqrt{p} & 0 \\
\vdots & \vdots \\
\sqrt{p} & 0 \\
\sqrt{r^2 / p} & \sqrt{q - r^2 / p} \\
\vdots & \vdots \\
\sqrt{r^2 / p} & \sqrt{q - r^2 / p}
\end{bmatrix}$$

::::

:::

## SBMs are GRDPGs

**Example 1** (cont'd): To perform community detection,

1. Note that $A$ is a GRDPG because $P = X I_{2, 0} X^\top$.
2. Compute the ASE $A \approx \hat{X} \hat{X}^\top$ with 
$\hat{X} = \hat{V} \hat{\Lambda}^{1/2}$.
3. Apply a clustering algorithm (e.g., $K$-means) to $\hat{X}$,  
noting that $\hat{X}$ approaches point masses as $n \to \infty$.

```{r, fig.height = 5, fig.width = 5, out.width = '50%', fig.cap = 'ASE of the adjacency matrix drawn from an SBM.'}
P.eigen = eigen(P)
X <- P.eigen$vectors[, 1:2] %*% diag(P.eigen$values[1:2] ** .5)

A.eigen <- eigen(A)
X.hat <- A.eigen$vectors[, 1:2] %*% diag(A.eigen$values[1:2] ** .5)
X.hat[, 2] <- -X.hat[, 2]
plot(X.hat, asp = 1, col = z * 2, xlab = NA, ylab = NA)
points(X, pch = 16, col = z * 2)
```

## DCBMs are GRDPGs

**Example 2**

* Convert example 1 to a DCBM by $P_{ij} \leftarrow \omega_i \omega_j P_{ij}$.
* This is equivalent to $P \leftarrow \Omega P \Omega$ where $\Omega = \diag(\omega_1, ..., \omega_n)$. 
* This is equivalent to left-multiplying the matrix of latent vectors $X$ by $\Omega$.
* The point masses in the latent structure of the SBM are "stretched out" toward the origin to form rays.

$$X = \begin{bmatrix}
\omega_1 \sqrt{p} & 0 \\ 
\vdots & \vdots \\ 
\omega_{n_1} \sqrt{p} & 0 \\ 
\omega_{n_1 + 1} \sqrt{r^2 / p} & \omega_{n_1 + 1} \sqrt{q - r^2 / p} \\ 
\vdots & \vdots \\
\omega_n \sqrt{r^2 / p} & \omega_n \sqrt{q - r^2 / p} \\ 
\end{bmatrix}$$

## DCBMs are GRDPGs

**Example 2** (cont'd): To perform community detection,

1. Note that $A$ is a GRDPG because $P = X I_{2, 0} X^\top$.
2. Compute the ASE $A \approx \hat{X} \hat{X}^\top$ with 
$\hat{X} = \hat{V} \hat{\Lambda}^{1/2}$.
3. Apply a clustering algorithm (e.g., kernel $K$-means with the cosine kernel) to $\hat{X}$, noting that $\hat{X}$ approaches rays as $n \to \infty$.

```{r fig.height = 2, fig.width = 6, out.width = '100%', fig.cap = 'DCBM (left) and its ASE (right).'}
set.seed(123)

n1 <- 2 ** 7
n2 <- 2 ** 7
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))
p <- 1/2
q <- 1/4
r <- 1/8
P <- matrix(r, nrow = n, ncol = n)
P[seq(n1), seq(n1)] <- p
P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q

omega <- runif(n)
P <- diag(omega) %*% P %*% diag(omega)
A <- draw.graph(P)

dcbm.plot <- ggraph(A, layout = 'auto') + 
  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
  geom_node_point(aes(colour = factor(z)), 
                  show.legend = FALSE) + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')

Xhat <- embedding(A, 2, 0)
Y <- embedding(P, 2, 0)

ase.plot <- ggplot() + 
  geom_point(aes(x = Xhat[, 1], 
                 y = Xhat[, 2], 
                 colour = factor(z)),
             size = .25) + 
  scale_colour_brewer(palette = 'Set1') + 
  labs(x = NULL, y = NULL) + 
  theme(legend.position = 'none') + 
  coord_fixed()

gridExtra::grid.arrange(dcbm.plot, ase.plot, nrow = 1)
```

## Example: UK Faculty Network

* Treating graphs with community structure as a GRDPG is useful for both EDA and inference. 
* The ASE of the UK faculty network (Nepusz et al., 2008) reveals three communities and that a DCBM is an appropriate model. 
* Clustering on the ASE results in low community detection error.

```{r, fig.width = 6, fig.height = 2, out.width = '100%', fig.cap = 'UK faculty network (left) and its ASE (right).'}
data(UKfaculty, package = 'igraphdata')
A <- as.matrix(igraph::as_adjacency_matrix(UKfaculty, type = 'both'))
A <- sign(A + t(A))
z <- factor(igraph::vertex_attr(UKfaculty)$Group)
A <- A[z != 4, z != 4]
z <- z[z != 4]
Xhat <- embedding(A, 3, 0)
n <- length(z)

graph.plot <- ggraph(A, layout = 'auto') + 
  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
  geom_node_point(aes(colour = factor(z)), 
                  show.legend = FALSE) + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')

ase.plot <- ggplot() + 
  geom_point(aes(x = Xhat[, 2], 
                 y = Xhat[, 3], 
                 colour = factor(z)),
             size = .5) + 
  scale_colour_brewer(palette = 'Set1') + 
  labs(x = 'comp 2', y = 'comp 3') + 
  theme(legend.position = 'none') + 
  coord_fixed()

gridExtra::grid.arrange(graph.plot, ase.plot, nrow = 1)
```

# PABMs are GRDPGs

## Popularity Vectors

**Lemma** (Noroozi, Rimal, and Pensky, 2020):

$A$ is sampled from a PABM if $P$ can be described as:

1. Let each $P^{(k \ell)}$ denote the $n_k \times n_l$ matrix of edge probabilities 
between communities $k$ and $l$. 
2. Organize popularity parameters as vectors 
$\lambda^{(k \ell)} \in \mathbb{R}^{n_k}$ 
such that $\lambda^{(k \ell)}_i = \lambda_{k_i l}$ is the popularity parameter 
of the $i$^th^ vertex of community $k$ towards community $l$. 
3. Each block can be decomposed as 
$P^{(k \ell)} = \lambda^{(k \ell)} (\lambda^{(lk)})^\top$.

This result reveals that $P$ is rank $K^2$ and has a linear latent structure.

## PABMs are GRDPGs

**Theorem** (KTT): $A \sim \PABM(\{\lambda_{ik}\}_K)$ is equivalent to 
$A \sim \GRDPG_{p, q}(X U)$ with

* $p = K (K + 1) / 2$, $q = K (K - 1) / 2$;
* $U$ is an orthogonal matrix;
* $X \in \mathbb{R}^{n \times K^2}$ is a block diagonal matrix 
composed of popularity vectors with each block corresponding to a community. 
* $P = (X U) I_{pq} (X U)^\top$.

$$X = \begin{bmatrix}
\Lambda^{(1)} & \cdots & 0 \\
0 & \ddots & 0 \\
0 & \cdots & \Lambda^{(K)}
\end{bmatrix} 
\in \mathbb{R}^{n \times K^2}$$

$$\Lambda^{(k)} = \begin{bmatrix} 
\lambda^{(k1)} & \cdots & \lambda^{(kK)} 
\end{bmatrix} 
\in \mathbb{R}^{n_k \times K}$$

$$A \sim \PABM(\{\lambda_{ik}\}_K) \text{ iff } A \sim \GRDPG_{p, q}(X U)$$

## Orthogonal Spectral Clustering

**Theorem** (KTT): 
If $P = V \Lambda V^\top$ and $B = n V V^\top$,  
then $B_{ij} = 0$ if $z_i \neq z_j$.

**Algorithm**: Orthogonal Spectral Clustering:

1. Let $V$ be the eigenvectors of $A$ corresponding to the $K (K+1)/2$ most 
positive and $K (K-1) / 2$ most negative eigenvalues.
2. Compute $B = |n V V^\top|$ applying $|\cdot|$ entry-wise.
3. Construct graph $G$ using $B$ as its similarity matrix.
4. Partition $G$ into $K$ disconnected subgraphs.

## Orthogonal Spectral Clustering

**Theorem** (KTT):  
Let $\hat{B}$ with entries $\hat{B}_{ij}$ be the affinity matrix from OSC. 
Then $\forall$ pairs $(i, j)$ belonging to different communities 
and sparsity factor satisfying $n \rho_n = \omega\big((\log n)^{4c}\big)$,  
$\max_{i, j} \hat{B}_{ij} = 
O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)$ 
as $n \to \infty$.

**Corollary**: OSC results in zero clustering error as $n \to \infty$, with probability 1.

## Sparse Subspace Clustering

**Corollary**: The ASE of $A \sim \PABM(\{\lambda^{(k \ell)}\}_K)$ lies near a collection of $K$-dimensional subspaces in $K^2$ dimensions.

**Algorithm**: Sparse Subspace Clustering (Elhamifar & Vidal, 2009):

1. Solve $n$ optimization problems $c_i = \arg\min_c \|c\|_1$ 
subject to $x_i = X^\top c$ and $c^{(i)} = 0$.  
This is typically performed via LASSO: 
$c_i = \arg\min \frac{1}{2} \|x_i - X_{-i}^\top c\|_2^2 + \lambda \|c\|_1$
2. Compile solutions $C = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix}$.
3. Construct affinity matrix $B = |C| + |C^\top|$.

## Sparse Subspace Clustering

Noroozi et al. observed that the rank of $P$ is $K^2$ 
and the columns of $P$ belonging to each community have rank $K$ 
to justify SSC for the PABM.

$$c_i = \arg\min_{c} \|c\|_1 \text{ subject to } A_{\cdot, i} = A c
\text{ and } c^{(i)} = 0$$

They were able to show that this obeys SDP if we replace $A$ with $P$.

GRDPG-based approach: Apply SSC to the ASE of $A$.

Stronger result: Apply SSC to the eigenvectors of $A$.

$$c_i = \arg\min_{c} \|c\|_1 \text{ subject to } 
\hat{v}_i = \hat{V}^\top c \text{ and } c^{(i)} = 0$$
$$A \approx \hat{V} \hat{\Lambda} \hat{V}^\top$$

## Sparse Subspace Clustering

**Theorem** (KTT):

Let

* $P_n$ describe the edge probability matrix of the PABM with $n$ vertices, and 
$A_n \sim \BG(P_n)$;
* $\hat{V}_n$ be the matrix of eigenvectors of $A_n$ corresponding to the 
$K (K + 1) / 2$ most positive and $K (K - 1) / 2$ most negative eigenvalues.

Then

* For some $\lambda > 0$ and $N < \infty$, $\sqrt{n} \hat{V}_n$ obeys the Subspace Detection Property with probability 1 when $n > N$.

Remarks:

* For large $n$, we can identify $\lambda$ for SDP (Wang and Xu, 2016).
* SDP does not guarantee community detection.

## Simulation Results

1. $z_1, ..., z_n \iid \Categorical(1 / K, ..., 1 / K)$
2. $\lambda_{ik} \stackrel{iid}{\sim} \Betadist(a_{ik}, b_{ik})$  
$a_{ik} = \begin{cases} 2 & z_i = k \\ 1 & z_i \neq k \end{cases}$, 
$b_{ik} = \begin{cases} 1 & z_i = k \\ 2 & z_i \neq k \end{cases}$
3. $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$
4. $A \sim \BG(P)$

```{r, fig.height = 2, fig.width = 8, out.width = '100%'}
clustering.df <- readr::read_csv('~/dev/pabm-grdpg/simulation-results/cluster-sim-balanced.csv')

clustering.df %>%
  dplyr::group_by(n, K) %>%
  dplyr::summarise(
    med.err = median(error.osc),
    first.q = quantile(error.osc, .25),
    third.q = quantile(error.osc, .75),
    med.err.ssc = median(error.ssc.ase),
    first.q.ssc = quantile(error.ssc.ase, .25),
    third.q.ssc = quantile(error.ssc.ase, .75),
    med.err.louvain = median(error.louvain),
    first.q.louvain = quantile(error.louvain, .25),
    third.q.louvain = quantile(error.louvain, .75),
    med.err.ssc.A = median(error.ssc.A),
    first.q.ssc.A = quantile(error.ssc.A, .25),
    third.q.ssc.A = quantile(error.ssc.A, .75)
  ) %>% 
  dplyr::ungroup() %>% 
  ggplot() +
  theme_bw() + 
  theme(legend.position = 'right',
        legend.direction = 'vertical',
        legend.justification = 0) + 
  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096),
                labels = c(7, 8, 9, 10, 11, 12)) +
  scale_y_log10() + 
  labs(y = 'error count', 
       x = expression(log[2] (n)),
       colour = NULL, shape = NULL) +
  geom_line(aes(x = n, y = med.err * n,
                colour = 'Orthogonal Spectral Clustering')) +
  geom_point(aes(x = n, y = med.err * n,
                 colour = 'Orthogonal Spectral Clustering', 
                 shape = 'Orthogonal Spectral Clustering')) +
  geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
                    colour = 'Orthogonal Spectral Clustering'), width = .1) + 
  geom_line(aes(x = n, y = med.err.ssc * n,
                colour = 'Sparse Subspace Clustering on ASE')) +
  geom_point(aes(x = n, y = med.err.ssc * n,
                 colour = 'Sparse Subspace Clustering on ASE', 
                 shape = 'Sparse Subspace Clustering on ASE')) +
  geom_errorbar(aes(x = n, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
                    colour = 'Sparse Subspace Clustering on ASE'), width = .1) +
  geom_line(aes(x = n, y = med.err.ssc.A * n,
                colour = 'Sparse Subspace Clustering on Adj. Matrix')) +
  geom_point(aes(x = n, y = med.err.ssc.A * n,
                 colour = 'Sparse Subspace Clustering on Adj. Matrix', 
                 shape = 'Sparse Subspace Clustering on Adj. Matrix')) +
  geom_errorbar(aes(x = n, ymin = first.q.ssc.A * n, ymax = third.q.ssc.A * n,
                    colour = 'Sparse Subspace Clustering on Adj. Matrix'), 
                width = .1) +
  geom_line(aes(x = n, y = med.err.louvain * n,
                colour = 'Modularity Maximization')) + 
  geom_point(aes(x = n, y = med.err.louvain * n,
                 colour = 'Modularity Maximization', 
                 shape = 'Modularity Maximization')) + 
  geom_errorbar(aes(x = n, ymin = first.q.louvain * n, ymax = third.q.louvain * n,
                    colour = 'Modularity Maximization'), width = .1) + 
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both')
```

## Simulation Results

1. $z_1, ..., z_n \iid \Categorical(\alpha_1, ..., \alpha_K)$ for $\alpha_k = \frac{k^{-1}}{\sum_{\ell=1}^K \ell^{-1}}$
2. $\lambda_{ik} \stackrel{iid}{\sim} \Betadist(a_{ik}, b_{ik})$  
$a_{ik} = \begin{cases} 2 & z_i = k \\ 1 & z_i \neq k \end{cases}$, 
$b_{ik} = \begin{cases} 1 & z_i = k \\ 2 & z_i \neq k \end{cases}$
3. $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$
4. $A \sim \BG(P)$

```{r, fig.height = 2, fig.width = 8, out.width = '100%'}
clustering.df <- readr::read_csv('~/dev/pabm-grdpg/simulation-results/cluster-sim-imbalanced.csv')

clustering.df %>%
  dplyr::group_by(n, K) %>%
  dplyr::summarise(
    med.err = median(error.osc),
    first.q = quantile(error.osc, .25),
    third.q = quantile(error.osc, .75),
    med.err.ssc = median(error.ssc.ase),
    first.q.ssc = quantile(error.ssc.ase, .25),
    third.q.ssc = quantile(error.ssc.ase, .75),
    med.err.louvain = median(error.louvain),
    first.q.louvain = quantile(error.louvain, .25),
    third.q.louvain = quantile(error.louvain, .75),
    med.err.ssc.A = median(error.ssc.A),
    first.q.ssc.A = quantile(error.ssc.A, .25),
    third.q.ssc.A = quantile(error.ssc.A, .75)
  ) %>% 
  dplyr::ungroup() %>% 
  ggplot() +
  theme_bw() + 
  theme(legend.position = 'right',
        legend.direction = 'vertical',
        legend.justification = 0) + 
  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096),
                labels = c(7, 8, 9, 10, 11, 12)) +
  scale_y_log10() + 
  labs(y = 'error count', 
       x = expression(log[2] (n)),
       colour = NULL, shape = NULL) +
  geom_line(aes(x = n, y = med.err * n,
                colour = 'Orthogonal Spectral Clustering')) +
  geom_point(aes(x = n, y = med.err * n,
                 colour = 'Orthogonal Spectral Clustering', 
                 shape = 'Orthogonal Spectral Clustering')) +
  geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
                    colour = 'Orthogonal Spectral Clustering'), width = .1) + 
  geom_line(aes(x = n, y = med.err.ssc * n,
                colour = 'Sparse Subspace Clustering on ASE')) +
  geom_point(aes(x = n, y = med.err.ssc * n,
                 colour = 'Sparse Subspace Clustering on ASE', 
                 shape = 'Sparse Subspace Clustering on ASE')) +
  geom_errorbar(aes(x = n, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
                    colour = 'Sparse Subspace Clustering on ASE'), width = .1) +
  geom_line(aes(x = n, y = med.err.ssc.A * n,
                colour = 'Sparse Subspace Clustering on Adj. Matrix')) +
  geom_point(aes(x = n, y = med.err.ssc.A * n,
                 colour = 'Sparse Subspace Clustering on Adj. Matrix', 
                 shape = 'Sparse Subspace Clustering on Adj. Matrix')) +
  geom_errorbar(aes(x = n, ymin = first.q.ssc.A * n, ymax = third.q.ssc.A * n,
                    colour = 'Sparse Subspace Clustering on Adj. Matrix'), 
                width = .1) +
  geom_line(aes(x = n, y = med.err.louvain * n,
                colour = 'Modularity Maximization')) + 
  geom_point(aes(x = n, y = med.err.louvain * n,
                 colour = 'Modularity Maximization', 
                 shape = 'Modularity Maximization')) + 
  geom_errorbar(aes(x = n, ymin = first.q.louvain * n, ymax = third.q.louvain * n,
                    colour = 'Modularity Maximization'), width = .1) + 
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both')
```

## Simulation Results

1. $z_1, ..., z_n \iid \Categorical(1 / K, ..., 1 / K)$
2. $\lambda_{ik} \stackrel{iid}{\sim} \Betadist(a_{ik}, b_{ik})$  
$a_{ik} = \begin{cases} 1 & z_i = k \\ 2 & z_i \neq k \end{cases}$, 
$b_{ik} = \begin{cases} 2 & z_i = k \\ 1 & z_i \neq k \end{cases}$
3. $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$
4. $A \sim \BG(P)$

```{r, fig.height = 2, fig.width = 8, out.width = '100%'}
clustering.df <- readr::read_csv('~/dev/pabm-grdpg/simulation-results/cluster-sim-disassortative.csv')

clustering.df %>%
  dplyr::group_by(n, K) %>%
  dplyr::summarise(
    med.err = median(error.osc),
    first.q = quantile(error.osc, .25),
    third.q = quantile(error.osc, .75),
    med.err.ssc = median(error.ssc.ase),
    first.q.ssc = quantile(error.ssc.ase, .25),
    third.q.ssc = quantile(error.ssc.ase, .75),
    med.err.louvain = median(error.louvain),
    first.q.louvain = quantile(error.louvain, .25),
    third.q.louvain = quantile(error.louvain, .75),
    med.err.ssc.A = median(error.ssc.A),
    first.q.ssc.A = quantile(error.ssc.A, .25),
    third.q.ssc.A = quantile(error.ssc.A, .75)
  ) %>% 
  dplyr::ungroup() %>% 
  ggplot() +
  theme_bw() + 
  theme(legend.position = 'right',
        legend.direction = 'vertical',
        legend.justification = 0) + 
  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096),
                labels = c(7, 8, 9, 10, 11, 12)) +
  scale_y_log10() + 
  labs(y = 'error count', 
       x = expression(log[2] (n)),
       colour = NULL, shape = NULL) +
  geom_line(aes(x = n, y = med.err * n,
                colour = 'Orthogonal Spectral Clustering')) +
  geom_point(aes(x = n, y = med.err * n,
                 colour = 'Orthogonal Spectral Clustering', 
                 shape = 'Orthogonal Spectral Clustering')) +
  geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
                    colour = 'Orthogonal Spectral Clustering'), width = .1) + 
  geom_line(aes(x = n, y = med.err.ssc * n,
                colour = 'Sparse Subspace Clustering on ASE')) +
  geom_point(aes(x = n, y = med.err.ssc * n,
                 colour = 'Sparse Subspace Clustering on ASE', 
                 shape = 'Sparse Subspace Clustering on ASE')) +
  geom_errorbar(aes(x = n, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
                    colour = 'Sparse Subspace Clustering on ASE'), width = .1) +
  geom_line(aes(x = n, y = med.err.ssc.A * n,
                colour = 'Sparse Subspace Clustering on Adj. Matrix')) +
  geom_point(aes(x = n, y = med.err.ssc.A * n,
                 colour = 'Sparse Subspace Clustering on Adj. Matrix', 
                 shape = 'Sparse Subspace Clustering on Adj. Matrix')) +
  geom_errorbar(aes(x = n, ymin = first.q.ssc.A * n, ymax = third.q.ssc.A * n,
                    colour = 'Sparse Subspace Clustering on Adj. Matrix'), 
                width = .1) +
  geom_line(aes(x = n, y = med.err.louvain * n,
                colour = 'Modularity Maximization')) + 
  geom_point(aes(x = n, y = med.err.louvain * n,
                 colour = 'Modularity Maximization', 
                 shape = 'Modularity Maximization')) + 
  geom_errorbar(aes(x = n, ymin = first.q.louvain * n, ymax = third.q.louvain * n,
                    colour = 'Modularity Maximization'), width = .1) + 
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both')
```

## Example

\begin{figure}[H]
{\centering \includegraphics[scale=.33]{/home/johnkoo/dev/pabm-grdpg/summary_files/figure-latex/mp-1}
}
\caption{Adjacency matrices of (from left to right) the British MPs Twitter network (Greene and Cunningham, 2013), Political Blogs network (Adamic and Glance, 2005), and DBLP network (Gao et al., 2009) after sorting by the clustering outputted by OSC.}\label{fig:mp}
\end{figure}

\begin{table}[H]
\centering
\begin{tabular}[t]{l|r|r|r}
\hline
Network & MM & SSC-ASE & OSC\\
\hline
British MPs & 0.003 & 0.012 & 0.006\\
\hline
Political Blogs & 0.050 & 0.187 & 0.062\\
\hline
DBLP & 0.028 & 0.072 & 0.059\\
\hline
\end{tabular}
\end{table}

## Example: Karantaka Villages (Banerjee et al., 2013)

\begin{figure}[H]
{\centering \includegraphics[scale=.33]{/home/johnkoo/dev/pabm-grdpg/summary_files/figure-latex/unnamed-chunk-7-1}
}
\caption{Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by OSC (left). The villages studied here are, from left to right, 12, 31, and 46.}\label{fig:households-figure}
\end{figure}

\begin{table}[H]
\centering
\begin{tabular}[t]{l|r|r|r}
\hline
Network & MM & SSC-ASE & OSC\\
\hline
Village 12 & 0.270 & 0.291 & 0.227\\
\hline
Village 31 & 0.125 & 0.059 & 0.051\\
\hline
Village 46 & 0.052 & 0.069 & 0.056\\
\hline
\end{tabular}
\end{table}

## PABMs are GRDPGs

* Published in JCGS.
* Code, data, and preprint drafts available at https://github.com/johneverettkoo/pabm-grdpg
* R package WIP at https://github.com/johneverettkoo/osc

## Overview of Block Models as GRDPGs

::: columns

:::: {.column width='67%'}

```{r, cache = TRUE, fig.width = 6, fig.height = 2, out.width = '100%'}
set.seed(123)

n <- 64
n1 <- n / 2
n2 <- n / 2
z <- c(rep(1, n1), rep(2, n2))
p <- 1/2
q <- 1/2
r <- 1/8
P <- matrix(r, nrow = n, ncol = n)
P[seq(n1), seq(n1)] <- p
P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
X <- embedding(P, 2, 0)
A <- draw.graph(P)
Xhat <- embedding(A, 2, 0)
A.graph <- tidygraph::as_tbl_graph(A, directed = FALSE)

sbm.graph.plot <- ggraph(A, layout = 'stress') + 
  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
  geom_node_point(aes(colour = factor(z)),
                  show.legend = FALSE) + 
  labs(colour = NULL, 
       title = 'SBM') + 
  # theme_graph() + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')

sbm.ase.plot <- ggplot() + 
  geom_point(aes(x = Xhat[, 2], y = Xhat[, 1],
                 colour = factor(z)),
             shape = 4) + 
  geom_point(aes(x = X[, 2], y = X[, 1], 
                 colour = factor(z))) + 
  # coord_fixed() +
  theme_bw() + 
  theme(legend.position = 'none') + 
  labs(x = NULL, y = NULL, colour = NULL, 
       title = 'Point Masses') + 
  scale_colour_brewer(palette = 'Set1')

gridExtra::grid.arrange(sbm.graph.plot, sbm.ase.plot, nrow = 1)
```

```{r, cache = TRUE, fig.width = 6, fig.height = 2, out.width = '100%'}
set.seed(123)

n <- 128
n1 <- n / 2
n2 <- n / 2
z <- c(rep(1, n1), rep(2, n2))
p <- 1/2
q <- 1/2
r <- 1/8
P <- matrix(r, nrow = n, ncol = n)
P[seq(n1), seq(n1)] <- p
P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
omega <- rbeta(n, 2, 1)
P <- P * (omega %*% t(omega))
X <- embedding(P, 2, 0)
A <- draw.graph(P)
Xhat <- embedding(A, 2, 0)
Xhat[, 2] <- -Xhat[, 2]
A.graph <- tidygraph::as_tbl_graph(A, directed = FALSE)

dcbm.graph.plot <- ggraph(A, layout = 'stress') + 
  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
  geom_node_point(aes(colour = factor(z)),
                  show.legend = FALSE) + 
  labs(colour = NULL, 
       title = 'DCBM') + 
  # theme_graph() + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')

dcbm.ase.plot <- ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2],
                 colour = factor(z)),
             shape = 4) + 
  geom_point(aes(x = X[, 1], y = X[, 2], 
                 colour = factor(z))) + 
  # coord_fixed() +
  theme_bw() + 
  theme(legend.position = 'none') + 
  labs(x = NULL, y = NULL, colour = NULL, 
       title = 'Rays') + 
  scale_colour_brewer(palette = 'Set1')

gridExtra::grid.arrange(dcbm.graph.plot, dcbm.ase.plot, nrow = 1)
```

```{r, cache = TRUE, fig.width = 6, fig.height = 2, out.width = '100%'}
set.seed(123)

n <- 64
Pz <- generate.P.beta(n)
P <- Pz$P
z <- Pz$clustering
X <- embedding(P, 3, 1)
A <- draw.graph(P)
Xhat <- embedding(A, 3, 1)
# Xhat[, 1] <- -Xhat[, 1]
A.graph <- tidygraph::as_tbl_graph(A, directed = FALSE)

pabm.graph.plot <- ggraph(A, layout = 'stress') + 
  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
  geom_node_point(aes(colour = factor(z)),
                  show.legend = FALSE) + 
  labs(colour = NULL, 
       title = 'PABM') + 
  # theme_graph() + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')

pabm.ase.plot <- ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2],
                 colour = factor(z)),
             shape = 4) + 
  geom_point(aes(x = X[, 1], y = X[, 2], 
                 colour = factor(z))) + 
  # coord_fixed() +
  theme_bw() + 
  theme(legend.position = 'none') + 
  labs(x = NULL, y = NULL, colour = NULL, 
       title = 'Subspaces (Projected)') + 
  scale_colour_brewer(palette = 'Set1')

gridExtra::grid.arrange(pabm.graph.plot, pabm.ase.plot, nrow = 1)
```

::::

:::: {.column width='33%'}

&nbsp;

&nbsp;

* K-means clustering
* Gaussian mixture models

&nbsp;

* K-means with cosine similarity
* GMM on angles

&nbsp;

&nbsp;

* Orthogonal spectral clustering
* Sparse subspace clustering

::::

:::

# Manifold Block Models

## Motivating Example

Macaque visuotactile brain areas and connections network  
(Negyessy et al., 2006)

* ASE suggests community structure.
* Structure is nonlinear---$K$-means, OSC, SSC, etc. are not appropriate for these data.

```{r macaque-graph, fig.cap = 'Macaque visuotactile brain areas and connections network (left) and ASE (right). The vertices represent brain areas.', fig.height = 2, fig.width = 6, out.width = '100%'}
data(macaque, package = 'igraphdata')
A <- as.matrix(igraph::as_adjacency_matrix(macaque, type = 'both'))
A <- sign(A + t(A))
z <- as.numeric(factor(igraph::vertex_attr(macaque)$shape))

graph.plot <- ggraph(A, layout = 'stress') + 
  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
  geom_node_point(aes(colour = factor(z)), 
                  show.legend = FALSE) + 
  labs(colour = NULL) + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')

Xhat <- embedding(A, 3, 0)

ase.plot <- ggplot() + 
  geom_point(aes(x = Xhat[, 1],
                 y = Xhat[, 2],
                 colour = factor(z))) + 
  labs(x = NULL, y = NULL) + 
  theme_bw() + 
  theme(legend.position = 'none') + 
  coord_fixed() + 
  scale_colour_brewer(palette = 'Set1')

gridExtra::grid.arrange(graph.plot, ase.plot, nrow = 1)
```

## Manifold Block Model

Let $p, q \geq 0$, $d = p + q \geq 1$, $1 \leq r < d$, $K \geq 2$, and $n > K$ be integers. 
Define manifolds $\mathcal{M}_1, ..., \mathcal{M}_K \in \mathcal{X}$ for $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1]\}$ each by continuous function $g_k : [0, 1]^r \to \mathcal{X}$. 
Define probability distribution $F$ with support $[0, 1]^r$. 
Then the following mixture model is a *manifold block model*:

\begin{enumerate}
\item Draw labels $z_1, ..., z_n \iid \Categorical(\alpha_1, ..., \alpha_K)$.
\item Draw latent vectors by first taking $t_1,..., t_n \iid F$ and then computing each $x_i = g_{z_i}(t_i)$. 
\item Compile the latent vectors into data matrix $X = [ x_1 \mid \cdots \mid x_n ]^\top$ and define the adjacency matrix as $A \sim \GRDPG_{p,q}(X)$. 
\end{enumerate}

## Manifold Block Model

```{r intersect-curves-example, fig.height = 2, fig.width = 8, fig.cap = 'Latent vectors on intersecting curves (left), along with an RDPG drawn from this configuration (center) and its ASE (right).', cache = TRUE, out.width='100%'}
f1 <- function(t) {
  x1 <- t ^ 2
  x2 <- 2 * t - 2 * t ^ 2
  return(cbind(x1, x2))
}

f2 <- function(t) {
  x1 <- 2 * t - 2 * t ^ 2
  x2 <- 1 - 2 * t + t ^ 2
  return(cbind(x1, x2))
}

set.seed(314159)

n1 <- 2 ** 7
n2 <- n1
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))

a <- 1
b <- 1
t1 <- rbeta(n1, a, b)
t2 <- rbeta(n2, a, b)
T <- rbind(t1, t2)

X1 <- f1(t1)
X2 <- f2(t2)
X <- rbind(X1, X2)

latent.vecs <- ggplot() + 
  geom_point(aes(x = X[, 1], y = X[, 2], colour = factor(z)), size = .5) + 
  # coord_fixed() + 
  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
  theme_bw() + 
  theme(legend.position = 'none') + 
  scale_colour_brewer(palette = 'Set1')

P <- X %*% t(X)
diag(P) <- 0
A <- draw.graph(P)

rdpg <- ggraph(A, layout = 'stress') + 
  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
  geom_node_point(aes(colour = factor(z)), 
                  show.legend = FALSE) + 
  labs(colour = NULL) + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')

Xhat <- embedding(A, 2, 0)

ase <- ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], colour = factor(z)), size = .5) + 
  # coord_fixed() + 
  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
  theme_bw() + 
  theme(legend.position = 'none') + 
  scale_colour_brewer(palette = 'Set1')

gridExtra::grid.arrange(latent.vecs, rdpg, ase, nrow = 1)
```

## $K$-Curves Clustering

\begin{algorithm}[H]
\label{alg:kcurves}
\scriptsize
\DontPrintSemicolon
\SetAlgoLined
\KwData{Adjacency matrix $A$, number of communities $K$, embedding dimensions $p$, $q$, stopping criterion $\epsilon$}
\KwResult{Community assignments $1, ..., K$, curves $g_1, ..., g_K$}
Compute $X$, the ASE of $A$ using the $p$ most positive and $q$ most negative eigenvalues and their corresponding eigenvectors.\;
Initialize community labels $z_1, ..., z_n$.\;
\Repeat {the change in $\sum_k \sum_{i \in C_k} \|x_i - g_k(t_i)\|^2$ is less than $\epsilon$} {
\For {$k = 1, ..., K$} {
Define $X_k$ as the rows of $X$ for which $z_i = k$.\;
Fit curve $g_k$ and positions $t_{k_i}$ to $X_k$ by minimizing $\sum_{k_i} \|x_{k_i} - g_k(t_{k_i})\|^2$.\;
}
\For {$k = 1, ..., K$} {
Assign $z_i \leftarrow \arg\min_\ell \|x_i - g_\ell(t_i)\|^2$.\;
}
}
\caption{$K$-curves clustering.}
\end{algorithm}

## $K$-Curves Clustering

```{r multifit, cache = TRUE, fig.height = 3, fig.width = 8, fig.cap = '$K$-curves clustering fits on the ASE using various approaches. Left: quadratic fit with random initialization. Middle: quadratic fit initialized via spectral clustering. Right: cubic fit initialized via spectral clustering.', out.width = '100%'}
f1 <- function(t) {
  x1 <- t ^ 2
  x2 <- 2 * t - 2 * t ^ 2
  return(cbind(x1, x2))
}

f2 <- function(t) {
  x1 <- 2 * t - 2 * t ^ 2
  x2 <- 1 - 2 * t + t ^ 2
  return(cbind(x1, x2))
}

set.seed(314159)

n1 <- 2 ** 7
n2 <- n1
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))

a <- 1
b <- 1
t1 <- rbeta(n1, a, b)
t2 <- rbeta(n2, a, b)
T <- rbind(t1, t2)

X1 <- f1(t1)
X2 <- f2(t2)
X <- rbind(X1, X2)

P <- X %*% t(X)
diag(P) <- 0
A <- draw.graph(P)

Xhat <- embedding(A, 2, 0)

maxit <- 2 ** 4
animation.dir <- '.'
random.starts <- 2
eps <- 1e-6
normalize <- TRUE
animate <- FALSE

set.seed(1)
random <- manifold.clustering(Xhat, 2, 
                              A = A, 
                              parallel = TRUE,
                              intercept = FALSE,
                              maxit = maxit, 
                              normalize = normalize,
                              eps = eps,
                              verbose = FALSE,
                              animate = animate,
                              animation.dir = animation.dir,
                              animation.title = 'random')
random.plot <- plot.estimated.curves(Xhat, random) + 
  labs(x = NULL, y = NULL, title = 'random initialization') + 
  scale_colour_brewer(palette = 'Set1')

set.seed(2)
spectral <- manifold.clustering(
  Xhat, 2, 
  A = A, 
  parallel = TRUE, 
  intercept = FALSE, 
  normalize = normalize, 
  maxit = maxit, 
  eps = eps,
  initialization = 'spectral',
  verbose = FALSE,
  animate = animate,
  animation.dir = animation.dir,
  animation.title = 'spectral')
spectral.plot <- plot.estimated.curves(Xhat, spectral) + 
  labs(x = NULL, y = NULL, title = 'spectral initialization') + 
  scale_colour_brewer(palette = 'Set1')

set.seed(3)
cubic <- manifold.clustering(
  Xhat, 2, 
  degree = 3, 
  A = A, 
  parallel = TRUE, 
  intercept = FALSE, 
  normalize = normalize, 
  maxit = maxit, 
  eps = eps,
  initialization = 'spectral',
  verbose = FALSE,
  animate = animate,
  animation.dir = animation.dir,
  animation.title = 'cubic')
cubic.plot <- plot.estimated.curves(Xhat, cubic) + 
  labs(x = NULL, y = NULL, title = 'cubic fit') + 
  scale_colour_brewer(palette = 'Set1')

gridExtra::grid.arrange(random.plot, spectral.plot, cubic.plot, nrow = 1)
```

```{r loss, cache = TRUE, fig.cap = 'Clustering loss vs. iteration for each run of $K$-curves clustering.', fig.width = 8, fig.height = 2, out.width = '100%'}
loss.df <- dplyr::bind_rows(
  dplyr::tibble(loss = random$loss, 
                `initialization and curve` = 'random, quadratic',
                iteration = seq_along(random$loss)),
  dplyr::tibble(loss = spectral$loss, 
                `initialization and curve` = 'spectral, quadratic',
                iteration = seq_along(spectral$loss)),
  dplyr::tibble(loss = cubic$loss, 
                `initialization and curve` = 'spectral, cubic',
                iteration = seq_along(cubic$loss))
) %>% 
  dplyr::mutate(`initialization and curve` = factor(`initialization and curve`,
                                        levels = c('random, quadratic',
                                                   'spectral, quadratic',
                                                   'spectral, cubic')))

ggplot(loss.df) + 
  geom_point(aes(x = iteration, y = loss, colour = `initialization and curve`)) + 
  geom_line(aes(x = iteration, y = loss, colour = `initialization and curve`)) + 
  # ylim(0, max.loss) +
  scale_y_log10() +
  labs(y = 'clustering loss') + 
  theme_bw()
```

## $K$-Curves Clustering

$$L(z_1, ..., z_n, g_1, ..., g_K; X) = \frac{1}{n} \sum_{k=1}^K \sum_{i: z_i = k} \|x_i - g_k(t_i)\|^2$$

**Theorem** (KTT):  
Let $A \sim \MBM(\{\alpha_1, ..., \alpha_K\}, \{F, ..., F\}, \{g_1, ..., g_K\}; \rho_n)$ such that $F$ has support $[0, 1]$, and each $g_k(t) = g(t; p_k)$ is is a curve of order $R$ that does not self-intersect (for any $s \neq t$, $g_k(s) \neq g_k(t)$). 
Suppose that for each community $k$, we have labels for at least $R + 1$ vertices. 
Then if $n \rho_n = \omega(\log^{4c} n)$, as $n \to \infty$, the estimates outputted by $K$-curves clustering are such that 
$$L(\hat{z}_1, ..., \hat{z}_n, \hat{g}_1, ..., \hat{g}_K; X) \stackrel{p}{\to} 0.$$

## Simulation Results

1. Draw $z_1, ..., z_n \iid \Multinomial(1/2, 1/2)$.
2. Draw $t_1, ..., t_n \iid \Uniform(0, 1)$. 
3. Let each $x_i = g_{z_i}(t_i)$ where  $g_1(t) = [ t^2, 2 t (1-t) ]^\top$ and $g_2(t) = [ 2 t (1-t), (1-t)^2 ]^\top$. 
Collect the latent vectors into matrix $X = [ x_1 \mid \cdots \mid x_n ]^\top$. 
4. Draw $A \sim \GRDPG_{2,0}(X)$ .

```{r sim-curves, fig.height = 3, fig.width = 8, cache = TRUE, out.width = '100%'}
n.vec <- 2 ^ c(7, 8, 9, 10, 11)
nsamp.vec <- c(0, 4, 8)

clustering.df <- readr::read_csv('~/dev/manifold-block-models/simulations/balanced-2/balanced.csv')

clust.summary.df <- clustering.df %>% 
  dplyr::group_by(n, nsamp) %>% 
  dplyr::summarise(
    med.count = median(error.count),
    first.q.count = quantile(error.count, .25),
    third.q.count = quantile(error.count, .75),
    med.rate = median(error.rate),
    first.q.rate = quantile(error.rate, .25),
    third.q.rate = quantile(error.rate, .75)
  ) %>% 
  dplyr::ungroup()

ggplot(clust.summary.df) + 
  theme_bw() + 
  theme(text = element_text(size = 10)) + 
  scale_y_log10() +
  scale_x_log10(breaks = n.vec) +
  labs(y = 'community detection error rate',
       x = 'n',
       colour = 'known labels\nper community', 
       shape = NULL) + 
  geom_line(aes(x = n, y = med.rate, colour = factor(nsamp))) + 
  geom_point(aes(x = n, y = med.rate, colour = factor(nsamp))) + 
  geom_errorbar(aes(x = n, 
                    ymin = first.q.rate, ymax = third.q.rate, 
                    colour = factor(nsamp)),
                width = .125) + 
  scale_colour_brewer(palette = 'Set1')
```

## Simulation Results

* Same setup as before but fix $n = 512$ and only random initializations.
* Repeat simulation 256 times and look at the distribution of fitted curves and loss values.

```{r repeat-curves, fig.height = 2, fig.width = 8, cache = TRUE, out.width = '100%'}
plot.repeat.study <- function(repeat.df) {
  repeat.df <- na.omit(repeat.df)
  min.t <- 0
  max.t <- 1
  length.out <- 1e2
  t. <- seq(min.t, max.t, length.out = length.out)
  niter <- nrow(repeat.df)
  p.df <- repeat.df[c('p121', 'p122', 'p131', 'p132', 
                      'p221', 'p222', 'p231', 'p232')]
  p.clust <- kmeans(p.df, 4)$cluster
  loss.clust <- kmeans(repeat.df$loss, 2)$cluster
  curves.df <- foreach::foreach(i = seq(niter), .combine = dplyr::bind_rows) %dopar% {
    p1 <- matrix(c(repeat.df$p121[i], repeat.df$p122[i], 
                   repeat.df$p131[i], repeat.df$p132[i]),
                 nrow = 2, byrow = TRUE)
    p1 <- rbind(0, p1)
    p2 <- matrix(c(repeat.df$p221[i], repeat.df$p222[i],
                   repeat.df$p231[i], repeat.df$p232[i]),
                 nrow = 2, byrow = TRUE)
    p2 <- rbind(0, p2)
    p <- list(p1, p2)
    plyr::ldply(seq_along(p), function(k) {
      curve <- bezier.curve(t., p[[k]])
      if (mean(curve[, 1]) < 0) {
        curve[, 1] <- -curve[, 1]
      }
      dplyr::tibble(t = t.,
                    x = curve[, 1],
                    y = curve[, 2],
                    z = k,
                    iter = i,
                    p.clust = p.clust[i],
                    loss.clust = loss.clust[i],
                    loss = repeat.df$loss[i])
    })
  }
  ggplot(curves.df) + 
    geom_path(aes(x = y, y = x, group = interaction(z, iter),
                  colour = loss),
              alpha = 1 / log(niter)) + 
    theme_bw() + 
    coord_fixed() + 
    labs(colour = 'loss',
         x = NULL,
         y = NULL) + 
    viridis::scale_color_viridis()
}

plot.repeat.study(repeat.df)
```

```{r repeat-losses, fig.height = 2, fig.width = 5, cache = TRUE, out.width = '67%'}
repeat.df <- readr::read_csv('~/dev/manifold-block-models/simulations/repeat-study/repeat-study.csv')
ggplot(repeat.df) + 
  geom_histogram(aes(x = loss),
                 colour = 'black', fill = 'white') + 
  labs(x = 'loss values') + 
  theme_bw()
```

## Simulation Results

1. Draw $z_1, ..., z_n \iid \Multinomial(1/3, 1/3, 1/3)$.
2. Draw $t_1, ..., t_n \iid \Uniform(0, 1)$. 
3. Let each $x_i = g_{z_i}(t_i)$ where
    i. $g_1(t) = [ 2 t (t-1), t^2, 0 ]^\top$
    ii. $g_2(t) = [ 0, t^2, 2 t (t-1) ]^\top$
    iii. $g_3(t) = [2 t (t-1), t^2, 2 t (t-1) ]^\top$.
4. Collect the latent vectors into matrix $X = [ x_1 \mid \cdots \mid x_n ]^\top$.
5. Draw $A \sim \GRDPG_{3,0}(X)$.

```{r three-curves, cache = TRUE, fig.height = 4, fig.width = 15, fig.cap = 'Latent vectors (left) and ASE (right) of one realization of the mixture model in the simulation setup with $K=3$ manifolds.', out.width = '100%'}
par(mfrow = c(1, 2))

n <- 1024

p.list <- list(matrix(c(0, 1, 0, 
                        0, 0, 1,
                        0, 0, 0),
                      nrow = 3, ncol = 3),
               matrix(c(0, 0, 0, 
                        0, 0, 1,
                        0, 1, 0),
                      nrow = 3, ncol = 3),
               matrix(c(0, 1, 0, 
                        0, 0, 1,
                        0, 1, 0),
                      nrow = 3, ncol = 3))

K <- length(p.list)

z <- sample(seq(K), n, replace = TRUE)
z <- sort(z)

t <- runif(n)

X <- lapply(seq(K), function(k) {
  bezier.curve(t[z == k], 
               p.list[[k]], 
               intercept = TRUE)
}) %>% 
  do.call('rbind', .)

p <- ncol(X)
q <- 0

P <- grdpg.edge.prob.matrix(X, p, q)
A <- draw.graph(P)
Xhat <- embedding(A, p, q)

scatterplot3d::scatterplot3d(X[, 1], X[, 2], X[, 3],
                             color = z, 
                             xlab = NA, ylab = NA, zlab = NA,
                             pch = '.', asp = 1,
                             tick.marks = FALSE)

scatterplot3d::scatterplot3d(Xhat[, 1], Xhat[, 2], Xhat[, 3], 
                             color = z, 
                             xlab = NA, ylab = NA, zlab = NA, 
                             pch = '.', asp = 1, 
                             tick.marks = FALSE,
                             angle = 40)
```

## Simulation Results

1. Draw $z_1, ..., z_n \iid \Multinomial(1/3, 1/3, 1/3)$.
2. Draw $t_1, ..., t_n \iid \Uniform(0, 1)$. 
3. Let each $x_i = g_{z_i}(t_i)$ where
    i. $g_1(t) = [ 2 t (t-1), t^2, 0 ]^\top$
    ii. $g_2(t) = [ 0, t^2, 2 t (t-1) ]^\top$
    iii. $g_3(t) = [2 t (t-1), t^2, 2 t (t-1) ]^\top$.
4. Collect the latent vectors into matrix $X = [ x_1 \mid \cdots \mid x_n ]^\top$.
5. Draw $A \sim \GRDPG_{3,0}(X)$.

```{r sim-curves-3, fig.height = 3, fig.width = 8, out.width = '100%', cache = TRUE}
n.vec <- 2 ^ c(7, 8, 9, 10, 11)
nsamp.vec <- c(0, 4, 8)

clustering.df <- readr::read_csv('~/dev/manifold-block-models/simulations/balanced-3/balanced-3.csv')

clust.summary.df <- clustering.df %>% 
  dplyr::group_by(n, nsamp) %>% 
  dplyr::summarise(
    med.count = median(error.count),
    first.q.count = quantile(error.count, .25),
    third.q.count = quantile(error.count, .75),
    med.rate = median(error.rate),
    first.q.rate = quantile(error.rate, .25),
    third.q.rate = quantile(error.rate, .75)
  ) %>% 
  dplyr::ungroup()

ggplot(clust.summary.df) + 
  theme_bw() + 
  theme(text = element_text(size = 10)) + 
  scale_y_log10() +
  scale_x_log10(breaks = n.vec) +
  labs(y = 'community detection error rate',
       x = 'n',
       colour = 'known labels\nper community', 
       shape = NULL) + 
  geom_line(aes(x = n, y = med.rate, colour = factor(nsamp))) + 
  geom_point(aes(x = n, y = med.rate, colour = factor(nsamp))) + 
  geom_errorbar(aes(x = n, 
                    ymin = first.q.rate, ymax = third.q.rate, 
                    colour = factor(nsamp)),
                width = .125) + 
  scale_colour_brewer(palette = 'Set1')
```

## Example: Macaque Brain Areas Network

```{r macaque-graph-fit, fig.cap = 'Macaque visuotactile brain areas and connections network (left), ASE (middle), and fitted curves (right). The vertices represent brain areas.', fig.height = 2, fig.width = 8, out.width = '100%', cache = TRUE}
data(macaque, package = 'igraphdata')
A <- as.matrix(igraph::as_adjacency_matrix(macaque, type = 'both'))
A <- sign(A + t(A))
z <- as.numeric(factor(igraph::vertex_attr(macaque)$shape))

graph.plot <- ggraph(A, layout = 'stress') + 
  geom_edge_link(alpha = 1 / sqrt(n) / 2) + 
  geom_node_point(aes(colour = factor(z)), 
                  show.legend = FALSE) + 
  labs(colour = NULL) + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')

Xhat <- embedding(A, 2, 0)

ase.plot <- ggplot() + 
  geom_point(aes(x = Xhat[, 1],
                 y = Xhat[, 2],
                 colour = factor(z))) + 
  labs(x = NULL, y = NULL) + 
  theme_bw() + 
  theme(legend.position = 'none') + 
  scale_colour_brewer(palette = 'Set1') + 
  coord_fixed()

set.seed(123)
z.spectral <- spectral.clustering(A)
macaque.out <- manifold.clustering(Xhat, 2, 
                                   # initialization = 'random',
                                   initialization = z.spectral,
                                   # initialization = z,
                                   intercept = FALSE, 
                                   parallel = TRUE, 
                                   verbose = FALSE, 
                                   animate = FALSE, 
                                   curve.init = 'x')

fit.plot <- plot.estimated.curves(Xhat, macaque.out) + 
  scale_colour_brewer(palette = 'Set1')

gridExtra::grid.arrange(graph.plot, ase.plot, fit.plot, nrow = 1)
```

## Example: Drosophila connectome (Eichler et al., 2017)

```{r mbconnectome-graph, fig.height = 2, figh.width = 6, cache = TRUE, out.width = '67%'}
library(mbstructure)
data(MBconnectome)

graph.r <- generate.graph(newrdat, vdf.right)
A.r <- as.matrix(igraph::as_adjacency_matrix(graph.r$g, type = 'both'))
z.r <- as.numeric(graph.r$vdf$type)
n.r <- length(z.r)
A.r <- sign(A.r + t(A.r))
diag(A.r) <- 0
graph.r <- ggraph(A.r, layout = 'stress') + 
  geom_edge_link(alpha = 1 / sqrt(n.r) / 2) + 
  geom_node_point(aes(colour = factor(z.r)), 
                  show.legend = FALSE) + 
  labs(colour = NULL) + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')

graph.l <- generate.graph(newldat, vdf.left)
A.l <- as.matrix(igraph::as_adjacency_matrix(graph.l$g, type = 'both'))
z.l <- as.numeric(graph.l$vdf$type)
n.l <- length(z.l)
A.l <- sign(A.l + t(A.l))
diag(A.l) <- 0
graph.l <- ggraph(A.l, layout = 'stress') + 
  geom_edge_link(alpha = 1 / sqrt(n.l) / 2) + 
  geom_node_point(aes(colour = factor(z.l)), 
                  show.legend = FALSE) + 
  labs(colour = NULL) + 
  theme_void() + 
  scale_colour_brewer(palette = 'Set1')

gridExtra::grid.arrange(graph.l, graph.r, nrow = 1)
```

```{r mbconnectome-ase, fig.height = 2, fig.width = 8, out.width = '100%'}
Xhat.l <- embedding(A.l, 2, 1)
ase.l <- ggplot() + 
  geom_point(aes(x = Xhat.l[, 1], y = Xhat.l[, 2],
                 colour = factor(z.l))) + 
  coord_fixed(.5) +
  theme_bw() + 
  theme(legend.position = 'none') +
  scale_colour_brewer(palette = 'Set1') + 
  labs(x = NULL, y = NULL, colour = NULL)

Xhat.r <- embedding(A.r, 2, 1)
ase.r <- ggplot() + 
  geom_point(aes(x = Xhat.r[, 1], y = Xhat.r[, 2],
                 colour = factor(z.r))) + 
  coord_fixed(.5) +
  theme_bw() + 
  theme(legend.position = 'none') +
  scale_colour_brewer(palette = 'Set1') + 
  labs(x = NULL, y = NULL, colour = NULL)

gridExtra::grid.arrange(ase.l, ase.r, nrow = 1)
```

```{r mbconnectome-kcurves, fig.height = 2, fig.width = 8, cache = TRUE, out.width = '100%', fig.cap = 'Left and right hemispheres of the Drosophila connectome (top), ASEs (middle), and fitted curves (bottom).'}
set.seed(123456)

z.init.r <- sample.points(z.r, 6)
z.init.l <- sample.points(z.l, 6)

mbconnectome.out.r <- manifold.clustering(Xhat.r, 
                                          length(unique(z.r)), 
                                          initialization = z.init.r,
                                          intercept = FALSE, 
                                          parallel = TRUE, 
                                          verbose = FALSE, 
                                          animate = FALSE, 
                                          curve.init = 'x')
fitted.r <- plot.estimated.curves(Xhat.r, mbconnectome.out.r) + 
  scale_colour_brewer(palette = 'Set1') + 
  coord_fixed(.5) +
  labs(x = NULL, y = NULL) + 
  ylim(min(Xhat.r[, 2]), max(Xhat.r[, 2])) + 
  xlim(min(Xhat.r[, 1]), max(Xhat.r[, 1]))

mbconnectome.out.l <- manifold.clustering(Xhat.l, 
                                          length(unique(z.l)), 
                                          initialization = z.init.l,
                                          intercept = FALSE, 
                                          parallel = TRUE, 
                                          verbose = FALSE, 
                                          animate = FALSE, 
                                          curve.init = 'x')
fitted.l <- plot.estimated.curves(Xhat.l, mbconnectome.out.l) + 
  scale_colour_brewer(palette = 'Set1') + 
  coord_fixed(.5) +
  labs(x = NULL, y = NULL) + 
  ylim(min(Xhat.l[, 2]), max(Xhat.l[, 2])) + 
  xlim(min(Xhat.l[, 1]), max(Xhat.l[, 1]))

gridExtra::grid.arrange(fitted.l, fitted.r, nrow = 1)
```

# Additional Slides

## Expected Degree and Sparsity Factor

Let $d_1, ..., d_n$ be the degrees of vertices $v_1, ..., v_n$. 
For $P$ defined by a block model or GRDPG, $E[d_i] = O(n)$. 
In most real networks, the degree does not scale linearly with the size of the network.

Introduce a sparsity factor $\rho_n$ and draw $A \sim \BG(\rho_n P)$.

* Sparse regime: $\lim\limits_{n \to \infty} \rho_n \to 0$
* Dense regime: $\rho_n = 1$

Most results still require $n \rho_n = \omega (\log^c n)$, i.e., 
expected degree still diverges, but at a logarithmic rate instead of linearly.

## Subspace Detection Property

If $X$ obeys the *Subspace Detection Property*, then $B$ is sparse such that 
$B_{ij} = 0$ for each $x_i$ and $x_j$ belonging to different subspaces.

::: columns

:::: column

```{r, cache = TRUE, fig.height = 3, fig.width = 4, out.width = '50%'}
set.seed(314159)

n <- 64

par(mar = rep(1, 4))

x1 <- runif(n, 1 / sqrt(50), 1)
y1 <- rep(0, n)
x2 <- rep(0, n)
y2 <- x1
z <- c(rep(1, n), rep(2, n))
X <- cbind(c(x1, x2), c(y1, y2))
X <- X + cbind(runif(n * 2, 0, .025),
               runif(n * 2, 0, .025))
X <- X %*% rbind(c(cos(pi / 9), sin(pi / 9)),
                 c(-sin(pi/ 9), cos(pi / 9)))
plot(X, col = z * 2, asp = 1, xlab = NA, ylab = NA)
```

::::

:::: column

```{r, fig.height = 3, fig.width = 5, out.width = '75%'}
Y <- t(X)
Y <- sweep(Y, 2, apply(Y, 2, function(y) sqrt(sum(y ** 2))), `/`)
Y[is.nan(Y)] <- 0
lambda <- .05
N <- ncol(Y)
B <- plyr::aaply(seq(N), 1, function(i) {
  y <- Y[, i]
  X <- Y[, -i]
  betahat <- glmnet::glmnet(X, y, lambda = lambda, intercept = FALSE) %>% 
    coef() %>% 
    as.numeric()
  if (i != N) {
    betahat <- c(betahat[seq(i)], 0, betahat[seq(i + 1, N)])
  } else {
    betahat <- c(betahat, 0)
  }
  betahat <- betahat[-1]
  return(betahat)
}, .parallel = TRUE) %>% 
  abs()
B <- sweep(B, 2, apply(B, 2, max), `/`)
B[is.nan(B)] <- 0
W <- B + t(B)

qgraph::qgraph(sign(W), layout = 'spring', groups = factor(z), legend = FALSE)
```

::::

:::

**Theorem** (Wang and Xu, 2016): 
If data matrix $X$ consists of points lying close to 
low-dimensional subspaces such that:

1. Each point's distance to its subspace is sufficiently small.
2. The points corresponding to each subspace cover a sufficient amount of 
that subspace.
3. The cosine of the angles between pairs of subspaces is sufficiently small.

Then $X$ obeys SDP.

## Parameter Estimation

1. Let $A^{(kl)}$ be the block of edges between communities $k$ and $l$. 
2. For each $k, l \in [K]$, do:
    i. Compute the SVD $A^{(kl)} = U \Sigma V^\top$.
    ii. Assign $u^{(kl)}$ and $v^{(kl)}$ as the first columns of $U$ and $V$. 
    Assign $\sigma^{(kl)} \leftarrow \Sigma_{11}^{1/2}$. 
    iii. Assign $\hat{\lambda}^{(kl)} \leftarrow \pm \sigma^{(kl)} u^{(kl)}$ and 
    $\hat{\lambda}^{(kl)} \leftarrow \pm \sigma^{(kl)} v^{(kl)}$.

**Theorem** (KTT):

$$\max_{k, l} \|\hat{\lambda}^{(kl)} - \lambda^{(kl)}\| =
O_P \Big( \frac{(\log n_k)^c}{\sqrt{n_k}} \Big)$$

## Indefinite Orthogonal Group

$$\mathbb{O}(p, q) = \{Q : Q I_{p, q} Q^\top = I_{p, q}\}$$

* $Q^\top Q \neq I$
* If $A \sim \GRDPG_{p, q}(X)$, then $A \sim \GRDPG_{p, q}(X Q)$
* $(Q x)^\top (Q y) = x^\top Q^\top Q y \neq x^\top y$
* $\|Q\| \neq 1$ $\implies$ $\|Q x - Q y\| \neq \|x - y\|$

## Fitting Bezier Curves to Noisy Data

Let $g(t; p) = \sum_{r=0}^R p^{(r)} \binom{R}{r} t^r (1-t)^{R-r}$ be a Bezier curve of order $R$ with coefficients $p = [p^{(0)} \mid \cdots \mid p^{(R)} ]^\top$.

Objective: minimize squared loss: $L(p, T; X) = \|X - T p\|^2$.

$T$ is the $n \times (R+1)$ matrix of index points, i.e., $T_{ir} = \binom{R}{r} t_i^r (1-t_i)^{R-r}$.

If we know $T$, we can fit $p$: $\hat{p} = (T^\top T)^{-1} T^\top X$.

If we know $p$, we can fit each $t_1, ..., t_n$ by finding the roots of $\dot{L}(t_i; p, x_i) = \bigg(\sum_{r=0}^R \binom{R}{r} (-1)^r c_r t^r \bigg) \bigg(\sum_{r=0}^{R-1} \binom{R-1}{r} (-1)^r c_{r+1} t^r \bigg)$, where $c_R = \sum_{r=0}^R (-1)^{R-r} \binom{R}{r} p^{(r)}$.

## ASE Clustering for Nonintersecting Communities

\begin{algorithm}[H]
\label{alg:ase-singlelink}
\scriptsize
\DontPrintSemicolon
\SetAlgoLined
\KwData{Adjacency matrix $A$, number of communities $K$, embedding dimensions $(p, q)$.}
\KwResult{Community assignments $z_1, ..., z_n \in \{1, ..., K\}$.}
Compute $\hat{X}$, the ASE of $A$ using the $p$ most positive and $q$ most negative eigenvalues and their corresponding eigenvectors.\;
Apply single linkage clustering with $K$ communities on $\hat{X}$.\;
\caption{Graph manifold clustering via single-linkage.}
\end{algorithm}

```{r two-curves, fig.cap = 'Manifold block model from two disjoint curves. The latent configuration is on the left, a random dot product graph drawn from the latent configuration is on the middle, and the ASE is on the right.', fig.height = 2, fig.width = 8, out.width='100%'}
set.seed(123456)
n1 <- 2 ** 5
n2 <- 2 ** 5
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))
u1 <- runif(n1)
u2 <- runif(n2)
x1 <- cos(pi / 3 * u1)
y1 <- sin(pi / 3 * u1)
x2 <- 1 + cos(pi / 3 * u2 + pi)
y2 <- 1 + sin(pi / 3 * u2 + pi)
data.matrix <- cbind(c(x1, x2), c(y1, y2))
A <- draw.graph(data.matrix %*% t(data.matrix))
Xhat <- embedding(A, 2, 0)

par(mfrow = c(1, 3))
# par(fig = c(0, .7, 0, 1))
plot(data.matrix, col = z * 2, asp = 1,
     xlab = NA, ylab = NA, 
     # xaxt = 'n', yaxt = 'n',
     xlim = c(-.1, 1.1), ylim = c(0, 1))
# ggplot() +
#   geom_point(aes(x = data.matrix[, 1], y = data.matrix[, 2], colour = factor(z))) +
#   labs(x = NULL, y = NULL) +
#   scale_colour_brewer(palette = 'Set1') +
#   theme(legend.position = 'none') +
#   coord_fixed()
# par(fig = c(.3, 1, 0, 1))
qgraph::qgraph(A, groups = factor(z), legend = FALSE)
plot(Xhat, col = z * 2, asp = 1, 
     # xaxt = 'n', yaxt = 'n', 
     xlab = NA, ylab = NA)
```

## ASE Clustering for Nonintersecting Communities

```{r fig.height = 2, fig.width = 8, out.width='100%'}
set.seed(123456)
n1 <- 2 ** 5
n2 <- 2 ** 5
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))
u1 <- runif(n1)
u2 <- runif(n2)
x1 <- cos(pi / 3 * u1)
y1 <- sin(pi / 3 * u1)
x2 <- 1 + cos(pi / 3 * u2 + pi)
y2 <- 1 + sin(pi / 3 * u2 + pi)
data.matrix <- cbind(c(x1, x2), c(y1, y2))
A <- draw.graph(data.matrix %*% t(data.matrix))
Xhat <- embedding(A, 2, 0)

par(mfrow = c(1, 3))
# par(fig = c(0, .7, 0, 1))
plot(data.matrix, col = z * 2, asp = 1,
     xlab = NA, ylab = NA, 
     # xaxt = 'n', yaxt = 'n',
     xlim = c(-.1, 1.1), ylim = c(0, 1))
# ggplot() +
#   geom_point(aes(x = data.matrix[, 1], y = data.matrix[, 2], colour = factor(z))) +
#   labs(x = NULL, y = NULL) +
#   scale_colour_brewer(palette = 'Set1') +
#   theme(legend.position = 'none') +
#   coord_fixed()
# par(fig = c(.3, 1, 0, 1))
qgraph::qgraph(A, groups = factor(z), legend = FALSE)
plot(Xhat, col = z * 2, asp = 1, 
     # xaxt = 'n', yaxt = 'n', 
     xlab = NA, ylab = NA)
```

**Theorem** (KTT): Let $A$ be a graph on $n$ vertices from the MBM model in wherein $n \rho_n = \omega(\log^{c}{n})$ for some $c > 1$ and the manifolds $\mathcal{M}_1, \dots, \mathcal{M}_{K}$ satisfy $\min_{k \neq \ell} \min_{x \in \mathcal{M}_k, y \in \mathcal{M}_{\ell}} \|x  - y\| \geq \delta$ for some $\delta > 0$ not depending on $n$ or $K$. 
Let $\hat{X}$ be the ASE of $A$ onto $\mathbb{R}^{d}$. 
Define $A_n(\eta)$ as the event that an $\eta$-neighborhood graph constructed from the ASE of $A$ consists of exactly $K$ disconnected subgraphs that map exactly to each manifold. 
Then for some $C > 0$ and any $\eta \in (0, C \delta)$, $\lim\limits_{n \to \infty} P(A_n(\eta)) = 1$.

## References

* Adamic, L. A. and Glance, N. (2005). The political blogosphere and the 2004 U.S. election: Divided they blog. In *Proceedings of the 3rd International Workshop on Link Discovery*, LinkKDD '05. 
* Banerjee, A., Chandrasekhar, A. G., Duflo, E., and Jackson, M. O. (2013). The diffusion of Microfinance. Harvard Dataverse. 
* Eichler, K., Li, F., Litwin-Kumar, A., Park, Y., Andrade, I., Schneider-Mizell, C. M., Saumweber, T., Huser, A., Eschbach, C., Gerber, B., Fetter, R. D., Truman, J. W., Priebe, C. E., Abbot, L. F., Thum, A. S., Zlatic, M., and Cardona, A. (2017). The complete connectome of a learning and memory center in an insect brain. *bioRxiv*.

## References

* Elhamifar, E. and Vidal, R. (2009). Sparse Subspace Clustering. *2009 IEEE Conference on Computer Vision and Pattern Recognition*. 
* Gao, J., Liang, F., Fan, W., Sun, Y., and Han, J. (2009). Graph-based consensus maximization among multiple supervised and unsupervised models. *Advances in Neural Information Processing Systems 22*. 
* Greene, D. and Cunningham, P. (2013). Producing a unified graph representation from multiple social network views. *CoRR*. 
* Karrer, B. and Newman, M. E. J. (2011). Stochastic blockmodels and community structure in networks. *Physical Review E*. 
* Koo, J., Tang, M., and Trosset, M. W. (2022). Popularity adjusted block models are generalized random dot product graphs. *Journal of Computational and Graphical Statistics*.

## References

* Lorrain, F., and White, H. C. (1971). Structural equivalence of individuals in social networks. *The Journal of Mathematical Society*. 
* Nepusz, T., Petroczi, A., Negyessy, L, and Bazso, F. (2008). Fuzzy communities and the concept of bridgeness in complex networks. *Phys. Rev. E*. 
* Noroozi, M. and Pensky, M. (2022). The hierarchy of block models. *Sankhya A*.
* Noroozi, M., Rimal, R., and Pensky, M. (2019). Estimation and clustering in the popularity adjusted block model. *Journal of the Royal Statistical Society: Series B (Statistical Methodology)*.

## References

* Negyessy, L., Nepusz, T., Kocsis, L., and Bazso, F. (2006). Prediction of the main cortical areas and connections involved in the tactile function of the visual cortex by network analysis. *European Journal of Neuroscience*. 
* Rubin-Delanchy, P., Cape, J., Tang, M., and Priebe, C. E. (2022). A statistical interpretation of spectral embedding: The generalised random dot product graph. *Journal of the Royal Statistical Society: Series B (Statistical Methodology)*. 
* Sengupta, S. and Chen, Y. (2018). A block model for node popularity in networks with community structure. *Journal of the Royal Statistical Society, Series B (Statistical Methodology)*. 
* Sussman, D. L., Tang, M., Fishkind, D. E., and Priebe, C. E. (2012). A consistent adjacency spectral embedding for stochastic blockmodel graphs. *Journal of the American Statistical Association*.

## References

* Wang, Y.-X. and Xu, H. (2016). Noisy sparse subspace clustering. *Journal of Machine Learning Research*.